### 2020-08-28 LCA-GB.Rmd
predict scenic quality for England and Scotland

---
title: "LCA-GB"
author: "Yi-Min Chang Chien"
date: "14/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(dplyr)
library(fasterize)
library(magrittr)
library(MASS)
library(raster)
library(sf)
library(tidyr)
library(xgboost)
library(yarrr)
# library(ggplot2)
# library(ggstance)
# library(GWmodel)
# library(kableExtra)
# library(knitr)
# library(openxlsx)
# library(tidyr)
```

```{r}
#### Data preparation
## Download the boundary of Wales
temp <- tempfile()
temp2 <- tempfile()
download.file("https://datashare.is.ed.ac.uk/bitstream/handle/10283/2410/Wales_boundary.zip", temp)
unzip(zipfile = temp, exdir = temp2)
"Wales boundary.shp" %>%
  file.path(temp2, .) %>%
  st_read(stringsAsFactors = FALSE) %>% 
  st_transform(crs = 27700) %>%
  st_combine() ->
  Wales
rm(list = c("temp", "temp2"))

load("/Users/Yi-Min/Rsession/ScenicOrNot/scenicness/Grid/Hexgonal Grid/5km.RData")
load("/Users/Yi-Min/Rsession/ScenicOrNot/scenicness/Grid/Hexgonal Grid/2km.RData")
st_crs(hex.5k) == st_crs(Wales) # checking the coordinate reference system
st_crs(hex.2k) == st_crs(Wales) # checking the coordinate reference system
hex.5km <- hex.5k[Wales, ]
hex.2km <- hex.2k[Wales, ]
```

# Data preparation/sampling
```{r}
# Rasterising LANDMAP data
VS <- read_sf("http://lle.gov.wales/catalogue/item/LandmapVisualSensory.json") %>%
  st_make_valid() %>%
  mutate(ScenicQuality = factor(VS_46, levels = c("Low", "Moderate", "High", "Outstanding")),
             Integrity = factor(VS_47, levels = c("Low", "Moderate", "High", "Outstanding")),
             Character = factor(VS_48, levels = c("Low", "Moderate", "High", "Outstanding")),
                Rarity = factor(VS_49, levels = c("Low", "Moderate", "High", "Outstanding"))) %>%
  dplyr::select(ScenicQuality, Integrity, Character, Rarity) %>%
  fasterize(raster(extent(.), resolution = 50, crs = 27700),
            field = "ScenicQuality")

# have a look
plot(VS)

hex.5km$SQ <- raster::extract(VS, as(hex.5km, "Spatial"), na.rm = FALSE, df = FALSE, factors = TRUE) %>%
  rapply(function(x) which.max(tabulate(x)), how = "replace") %>%
  lapply(function(x) ifelse(is.null(x), NA, x)) %>%
  unlist() %>%
  factor(levels = c(1, 2, 3, 4)) %>%
  plyr::mapvalues(from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding"))

hex.2km$SQ <- raster::extract(VS, as(hex.2km, "Spatial"), na.rm = FALSE, df = FALSE, factors = TRUE) %>%
  rapply(function(x) which.max(tabulate(x)), how = "replace") %>%
  lapply(function(x) ifelse(is.null(x), NA, x)) %>%
  unlist() %>%
  factor(levels = c(1, 2, 3, 4)) %>%
  plyr::mapvalues(from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding"))
```

# Exploratory analysis (using Pirate plot)

```{r}
# Set margins of document for our new custom figure
par(mfrow = c(1, 1), mar = c(4,4.5,2,2))

# Create a custom pirate plot! (yar!)
bespoke_plot <- function(observation, formula, title, x.label, y.label){
  pirateplot(formula = formula,
             data = observation,
             pal = c("#EA4335", "#FBBC05","#34A853","#4285F4"),
             xlab = x.label,
             ylab = y.label,
             main = title,
             theme = 0, # set theme to 0
             bean.f.o = .4, # Bean fill
             point.o = .4, # Points
             inf.f.o = .4, # Inference fill
             inf.b.o = .8, # Inference border
             avg.line.o = 0.8, # Average line
             point.cex = .2, # Points
             quant = c(.1, .9)) #, # Adjust quantiles
             #ylim = c(0, 5))  # Adjust y axis limits
  text(x = c(1, 2, 3, 4), 
       y = 1 , 
       labels = paste(summary(observation$SQ)[1:4]), 
       cex = 1) #print number of rows of data
}

p1 <- bespoke_plot(observation = hex.5km, 
                   formula = Sce ~ SQ,
                   title = "Relationship between Scenic Quality and Scenicness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Scenic-Or-Not ratings")
p2 <- bespoke_plot(observation = hex.2km, 
                   formula = Sce ~ SQ,
                   title = "Relationship between Scenic Quality and Scenicness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Scenic-Or-Not ratings")

p3 <- bespoke_plot(observation = hex.5km, 
                   formula = Abs ~ SQ,
                   title = "Relationship between Scenic Quality and Absence",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Absence of human artefacts")
p4 <- bespoke_plot(observation = hex.2km, 
                   formula = Abs ~ SQ,
                   title = "Relationship between Scenic Quality and Absence",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Absence of human artefacts")

p5 <- bespoke_plot(observation = hex.5km, 
                   formula = Nat ~ SQ,
                   title = "Relationship between Scenic Quality and Naturalness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Naturalness of perceived land cover")
p6 <- bespoke_plot(observation = hex.2km, 
                   formula = Nat ~ SQ,
                   title = "Relationship between Scenic Quality and Naturalness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Naturalness of perceived land cover")

p7 <- bespoke_plot(observation = hex.5km, 
                   formula = Rem ~ SQ,
                   title = "Relationship between Scenic Quality and Remoteness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Remoteness from mechanised roads")
p8 <- bespoke_plot(observation = hex.2km, 
                   formula = Rem ~ SQ,
                   title = "Relationship between Scenic Quality and Remoteness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Remoteness from mechanised roads")

p9 <- bespoke_plot(observation = hex.5km, 
                   formula = Rug ~ SQ,
                   title = "Relationship between Scenic Quality and Ruggedness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Ruggedness of terrain")
p10 <- bespoke_plot(observation = hex.2km, 
                   formula = Rug ~ SQ,
                   title = "Relationship between Scenic Quality and Ruggedness",
                   x.label = "LANDMAP evaluation of scenc quality",
                   y.label = "Ruggedness of terrain")
```
# Ordinal regression
#
```{r}
##### 3.2. Include wildness components as additional predictors
model1 = polr(SQ ~ scale(Sce), data = as(hex.5km, "Spatial"), Hess = TRUE)
model2 = polr(SQ ~ scale(Sce) + scale(Abs), data = as(hex.5km, "Spatial"), Hess = TRUE)
model3 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat), data = as(hex.5km, "Spatial"), Hess = TRUE)
model4 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem), as(hex.5km, "Spatial"), Hess = TRUE)
model5 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem) + scale(Rug), as(hex.5km, "Spatial"), Hess = TRUE)
model6 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rug), as(hex.5km, "Spatial"), Hess = TRUE)

summary(model1)
summary(model2)
summary(model3)
summary(model4)
summary(model5)
summary(model6)

AIC.scores <- c(AIC(model1), AIC(model2), AIC(model3), AIC(model4), AIC(model5), AIC(model6))
names(AIC.scores) <- c("model1", "model2", "model3", "model4", "model5", "model6")
AICw <- geiger::aicw(AIC.scores)

table <- function(model){
  summary_table <- coef(model) %>% exp() %>% round(3) %>% as.data.frame() 
  pval <- 
    coef(summary(model))[1:length(model$coefficients), "t value"] %>% 
    abs() %>%
    pnorm(., lower.tail = FALSE)*2 
  tab <- cbind("Odds Ratio" = summary_table, "p value" = round(pval, 3))
  return(tab)
}

table(model1)
table(model2)
table(model3)
table(model4)
table(model5)
table(model6)
```
```{r}
##### 3.2. Include wildness components as additional predictors
model1 = polr(SQ ~ scale(Sce), data = as(hex.2km, "Spatial"), Hess = TRUE)
model2 = polr(SQ ~ scale(Sce) + scale(Abs), data = as(hex.2km, "Spatial"), Hess = TRUE)
model3 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat), data = as(hex.2km, "Spatial"), Hess = TRUE)
model4 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem), as(hex.2km, "Spatial"), Hess = TRUE)
model5 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem) + scale(Rug), as(hex.2km, "Spatial"), Hess = TRUE)
model6 = polr(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rug), as(hex.2km, "Spatial"), Hess = TRUE)

summary(model1)
summary(model2)
summary(model3)
summary(model4)
summary(model5)
summary(model6)

AIC.scores <- c(AIC(model1), AIC(model2), AIC(model3), AIC(model4), AIC(model5), AIC(model6))
names(AIC.scores) <- c("model1", "model2", "model3", "model4", "model5", "model6")
AICw <- geiger::aicw(AIC.scores)

table <- function(model){
  summary_table <- coef(model) %>% exp() %>% round(3) %>% as.data.frame() 
  pval <- 
    coef(summary(model))[1:length(model$coefficients), "t value"] %>% 
    abs() %>%
    pnorm(., lower.tail = FALSE)*2 
  tab <- cbind("Odds Ratio" = summary_table, "p value" = round(pval, 3))
  return(tab)
}

table(model1)
table(model2)
table(model3)
table(model4)
table(model5)
table(model6)
```

# GLM
```{r}
##### 3.1. Global Model: Generalised Linear Model
# hex.5km %>%
#   drop_na() %>%
#   as("Spatial") %>%
#   glm(SQ ~ scale(Sce) + scale(Abs) + scale(Nat) + scale(Rem) + scale(Rug),
#       data = .,
#       family = "poisson") %>%
#   summary()
# 
# R2 = 1 - glm$deviance/glm$null.deviance
```

# Machine learning - Gradient Boosted 
```{r}
# GBM hyper-parameters tuning
# From kaggle: https://www.sciencedirect.com/science/article/pii/S0378778817320844
##### Split data into training and validation dataset
set.seed(2046)
TrainIndex <- hex.2km %>% drop_na() %>% .$SQ %>% createDataPartition(p = .8, list = FALSE, times = 1)
Train <- hex.2km[TrainIndex, ]
Test  <- hex.2km[-TrainIndex, ]

nrounds = 1000
tune_grid <- expand.grid(
  nrounds = seq(from = 100, to = nrounds, by = 100),
  eta = c(0.05, 0.01, 0.005, 0.001),
  max_depth = 5,
  colsample_bytree = 1, # percent of columns to sample from for each tree
  min_child_weight = 1, # minimum node size
  subsample = 0.5,      # percent of training data to sample for each tree
  #gamma = 0
  gamma = 1
  # lambda = c(0, 1e-2, 0.1, 1, 100, 1000, 10000),
  # alpha = c(0, 1e-2, 0.1, 1, 100, 1000, 10000)
)

tune_control <- caret::trainControl(
  method = "cv", # cross-validation
  #method = "repeatedcv",
  #repeats = 2,
  number = 20, # with n folds 
  #index = createFolds(tr_treated$Id_clean), # fix the folds
  verboseIter = FALSE, # no training log
  allowParallel = FALSE, # FALSE for reproducible results
  returnData = FALSE,
  #classProbs = TRUE,
  #summaryFunction = multiClassSummary
  #summaryFunction = twoClassSummary
)

train_time <- system.time({
  xgb_tune <- caret::train(
    x = Train %>% dplyr::select(Sce, Abs, Nat, Rem, Rug) %>% st_drop_geometry() %>% as.matrix() %>% xgb.DMatrix(),
    y = Train$SQ,
    trControl = tune_control,
    tuneGrid = tune_grid,
    metric = "Accuracy",
    method = "xgbTree",
    # selectionFunction = "best", "oneSE","tolerance",
    # objective = "multi:softprob",
    # eval_metric = "mlogloss",
    num_class = 4
    )
  })
xgb_tune$bestTune
max(xgb_tune$results$Accuracy)

# helper function for the plots
tuneplot <- function(x) {
  ggplot(x) +
    coord_cartesian(ylim = c(max(x$results$Accuracy), min(x$results$Accuracy))) +
    theme_bw()
}

tuneplot(xgb_tune)

# Reference
# https://github.com/topepo/caret/issues/389
```
```{r}
# Model evaluation
#predicted = predict(xgb_tune, x_test)
# Predict outcomes with the test data
Test %>% dplyr::select(Sce, Abs, Nat, Rem, Rug) %>%
  st_drop_geometry() %>% 
  as.matrix() %>% 
  xgb.DMatrix() %>%
  predict(xgb_tune, ., reshape = T) %>%
  plyr::mapvalues(from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")) %>%
  data.frame(predict = .) %>%
  data.frame(actual = plyr::mapvalues(Test$SQ, from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")),
             predict = .) ->
  result

# confusion matrix
confusionMatrix(result$actual, result$predict) %>% print()

# Reference
# https://datascienceplus.com/extreme-gradient-boosting-with-r/

hex.2k$SQ.pred <-
  hex.2k %>%
  st_drop_geometry() %>%
  dplyr::select(Sce, Abs, Nat, Rem, Rug) %>%
  predict(xgb_tune, ., reshape = T) %>%
  plyr::mapvalues(from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")) %>%
  as.factor()
```
# Visualisation
```{r}
theme_map <- function(...) {
  theme_minimal() +
    theme(
      text = element_text(family = "Arial", color = "#22211d"), #"Ubuntu Regular"
      axis.line = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      # panel.grid.minor = element_line(color = "#ebebe5", size = 0.2),
      panel.grid.major = element_blank(), #element_line(color = "#ebebe5", size = 0.2),
      panel.grid.minor = element_blank(),
      plot.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA), 
      panel.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA), 
      legend.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA),
      panel.border = element_blank(),
      ...
    )
}

p <- ggplot(hex.2k) +
  # municipality polygons
  geom_sf(aes(fill = SQ.pred), color = NA) +
  theme_map() +
  labs(x = NULL, y = NULL, 
       title = "Predicted scenic quality") +
  theme(legend.position = "bottom",
        title =element_text(size = 15)) +
  scale_fill_manual(
  # in manual scales, one has to define colors, well, manually
  # I can directly access them using viridis' magma-function
  values = RColorBrewer::brewer.pal(4, "YlOrBr"), #rev(magma(9)),
  #breaks = brks_scale,
  name = "Quality levels",
  drop = FALSE,
  labels = c("Low", "Moderate", "High", "Outstanding"),
  guide = guide_legend(
    direction = "horizontal",
    keyheight = unit(6, units = "mm"),
    keywidth = unit(30 / length(labels), units = "mm"),
    title.position = 'top',
    title.theme = element_text(size = 15),
    # I shift the labels around, the should be placed 
    # exactly at the right end of each legend key
    # title.hjust = -0.9,
    label.hjust = 0.2,
    label.theme = element_text(size = 12),
    nrow = 1,
    byrow = T,
    # also the guide needs to be reversed
    reverse = F,
    label.position = "bottom")
  )

plot(test[, "SQ"])
```

### 2022-06-27 LCA-GB.Rmd
combine r and python code in RMarkdown

@@
 # Data preparation/sampling
+
 ```{r}
@@
 ```
+
 # Ordinal regression
+
 #
+
 ```{r}
@@
 ```
+
 ```{r}
@@
 # GLM
+
 ```{r}
@@
 # Machine learning - Gradient Boosted
+
 ```{r}
@@
 ```
+
 # Visualisation
+
 ```{r}

### 2022-06-27 paper3.Rmd
combine r and python code in RMarkdown

---
title: "paper3"
author: "Yi-Min Chang Chien"
date: '2022-06-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Package names
packages <- c("curl",
              "dplyr",
              "entropy",
              "magrittr",
              "raster",
              "readr",
              "sf",
              "stringr")

# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}

# Packages loading
invisible(lapply(packages, library, character.only = TRUE))

rm(list = c("installed_packages", "packages"))
```

```{r}
#### Data preparation
## Download the boundary of Wales
temp <- tempfile()
temp2 <- tempfile()
download.file("https://datashare.is.ed.ac.uk/bitstream/handle/10283/2410/Wales_boundary.zip", temp)
unzip(zipfile = temp, exdir = temp2)
"Wales boundary.shp" %>%
  file.path(temp2, .) %>%
  st_read(stringsAsFactors = FALSE) %>% 
  st_transform(crs = 27700) %>%
  st_combine() ->
  Wales
rm(list = c("temp", "temp2"))

## Download the visual and sensory aspect of the LANDMAP data  
VS <- read_sf("http://lle.gov.wales/catalogue/item/LandmapVisualSensory.json") %>%
  st_make_valid() %>%
  mutate(ScenicQuality = factor(VS_46, levels = c("Low", "Moderate", "High", "Outstanding")),
             Integrity = factor(VS_47, levels = c("Low", "Moderate", "High", "Outstanding")),
             Character = factor(VS_48, levels = c("Low", "Moderate", "High", "Outstanding")),
                Rarity = factor(VS_49, levels = c("Low", "Moderate", "High", "Outstanding")))

## Download the wildness measure and its components
#"https://github.com/yiminchangchien/scenicness/blob/2a3f9b996c11a7c480867524664401dbea9aea2b/Wildness.RData" %>% url() %>%

load("C:\\Users\\00631\\Downloads\\James\\Wildness.RData")

## Download Scenic-Or-Not data
sc <- read_tsv("http://scenicornot.datasciencelab.co.uk/votes.tsv",
               col_types = cols("ID" = col_number(),
                                "Lat" = col_double(),
                                "Lon" = col_double(),
                                "Average" = col_double(),
                                "Variance" = col_double(),
                                "Votes" = col_character(),
                                "Geograph URI" = col_character())) %>%
  rowwise() %>%
  mutate(Mean = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
         Variance = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% var(),
         Entropy = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
         Number = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% na.omit() %>% length())%>%
  st_as_sf(coords = c("Lon","Lat"), crs = 4326) %>%
  st_transform(crs = 27700)
```

```{r}
LANDMAP <- sc %>% 
  st_join(VS, .) %>% 
  aggregate(Votes~UID, data = ., paste0, collapse = ",") %>%
  merge(VS, ., by = "UID", all = T) %>%
  as_tibble() %>% 
  st_as_sf() %>%
  rowwise() %>%
  mutate(Scenicness.median = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
         Scenicness.mean = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
         Scenicness.entropy = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
         Scenicness.number = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% na.omit() %>% length(),
    Consultant = ifelse(str_detect(VS_1a, "Bronwen Thomas"), "B", "A")) %>%
  dplyr::select(UID, CLS_1, CLS_2, CLS_3, CLS_4, ScenicQuality, Consultant, Votes, Scenicness.median, Scenicness.mean, Scenicness.entropy, Scenicness.number) %>%
  ungroup()

plot(LANDMAP["ScenicQuality"], border = 'NA', key.pos = 1)  
```

Add wildness measures

```{r}
library(exactextractr)

Wildness <- raster('Z:\\R\\R projects\\LCA-GB\\James\\mce_eq256')

LANDMAP <- st_transform(LANDMAP, crs = st_crs(Wildness)) %>%
  exact_extract(Wildness, ., fun = c('mean', 'median', 'coefficient_of_variation', 'variance', 'quantile'), quantiles = c(0.25, 0.75)) %>%
  mutate(Wildness.iqr = q75 - q25) %>%
  dplyr::select(Wildness.mean = mean, Wildness.median = median, Wildness.iqr, Wildness.cv = coefficient_of_variation, Wildness.variance = variance) %>%
  bind_cols(LANDMAP, .) %>%
  st_transform(crs = 27700)

landmap <- st_cast(LANDMAP, "POLYGON")
```

```{r}
library(reticulate)

```

```{python}

import geopandas as gpd
import pandas as pd
from sklearn.preprocessing import OrdinalEncoder
from sklearn.compose import make_column_transformer

# landmap = geopandas.read_file("/Users/mac/R/LCA-GB/Data/Input/LANDMAP.geojson")
landmap = r.landmap
landmap = landmap.dropna(subset=['ScenicQuality'])

print(landmap)
print(landmap.describe)
print(landmap['ScenicQuality'].value_counts())

landmap.isnull().sum()

landmap_A = landmap.query("Consultant in ['A']")
landmap_B = landmap.query("Consultant in ['B']")

print(landmap.shape)
print(landmap_A.shape)
print(landmap_B.shape)

# def split_consultant(landmap):
#   X = landmap.loc[:, ['Scenicness.mean', 'Scenicness.entropy', 'Wildness.median', 'Wildness.iqr']]
#   y = landmap.loc[:, ['ScenicQuality']]
# 
#   # Instantiate the label encoder
#   transformer = make_column_transformer(
#     (OrdinalEncoder(categories=[['Low', 'Moderate', 'High', 'Outstanding']]), ['ScenicQuality'])
#   )
# 
#   y = transformer.fit_transform(y)
#   
#   y = pd.DataFrame(y, columns = ['ScenicQuality'])
#   return X, y
# 
# def EDA_plot(y, X):
#   import pandas as pd
#   # concatenating y and X along columns
#   landmap_concat = pd.DataFrame(pd.concat([y.reset_index(drop=True), X.reset_index(drop=True)], axis=1))
#   print(landmap_concat.shape)
# 
#   import seaborn as sns
#   sns.set()
#   sns.pairplot(landmap_concat, hue='ScenicQuality', palette='vlag', height=3)

# split train/test set
X = landmap.loc[:, ['Scenicness.mean']]
y = landmap.loc[:, ['ScenicQuality']]

# from sklearn.preprocessing import OrdinalEncoder
# OE = OrdinalEncoder()
# y[['Scale']] = OE.fit_transform(y[['ScenicQuality']])

scale_mapper = {'Low': int(0), 'Moderate': int(1), 'High': int(2), 'Outstanding': int(3)} 
y[['ScenicQuality']] = y[['ScenicQuality']].replace(scale_mapper).astype('int')

# X, y = split_consultant(landmap)
print(y.shape, X.shape)
# EDA_plot(y, X)

ratio = 0.8
from sklearn.model_selection import train_test_split
# Create train/test split
X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=1 - ratio, train_size=ratio, random_state=19, stratify=y)
```
## Hyperparameter Tuning
```{python}
%%time
from hyperopt import fmin, tpe, hp, STATUS_OK, Trials, space_eval

# Choose hyperparameter domain to search over
space = {

         'max_depth':hp.choice('max_depth', np.arange(1, 30, 1, dtype=int)),
  'colsample_bytree':hp.quniform('colsample_bytree', 0.3, 1.01, 0.05),
  'min_child_weight':hp.choice('min_child_weight', np.arange(1, 30, 1, dtype=int)),
         'subsample':hp.quniform('subsample', 0.3, 1.01, 0.05),
     'learning_rate':hp.choice('learning_rate',    np.arange(0.05, 1.01, 0.05)),
             'gamma': hp.quniform('gamma', 0.1, 5, 0.05),
         # 'objective':'reg:squarederror',
         'objective': 'multi:softmax',
       'eval_metric': 'rmse',
    }

def score(params, n_folds=5):

    #Cross-validation
    d_train = xgboost.DMatrix(X_coords,y)

    cv_results = xgboost.cv(params, d_train, nfold = n_folds, num_boost_round=500,
                        early_stopping_rounds = 10, metrics = 'rmse', seed = 0)

    loss = min(cv_results['test-rmse-mean'])

    return loss

def optimize(trials, space):

    best = fmin(score, space, algo=tpe.suggest, max_evals=2000,
                rstate=np.random.RandomState(333))#Add seed to fmin function
    return best
```

## Fit xgboost model

```{python}
import xgboost as xgb
# Specify sufficient boosting iterations to reach a minimum
num_round = 1000 #209.91

# Leave most parameters as default
param = {'objective': 'multi:softmax',  # Specify multiclass classification
         'num_class': 4,  # Number of possible output classes
         # 'tree_method': 'gpu_hist',  # Use GPU accelerated algorithm
         'eta': 0.451,
         'max_depth': 12,
         'gamma': 0,
         'subsample': 0.758,
         'colsample_bytree': 1,
         'min_child_weight': 0.2,
         'tree_method': 'hist',
         'predictor': 'cpu_predictor'
         }

# Convert input data from numpy to XGBoost format
dtrain = xgb.DMatrix(X_train, label=y_train)
dtest = xgb.DMatrix(X_test, label=y_test)

import time
gpu_res = {}  # Store accuracy result
start = time.time()

# Train model
# model = xgb.train(param, dtrain, num_round, evals=[(dtest, 'test')], evals_result=gpu_res)
# print("GPU Training Time: %s seconds" % (str(time.time() - start)))
model = xgb.train(param, dtrain, num_round, evals=[(dtest, 'test')])
print(time.time() - start)

# Evaluate model using test data
from sklearn.metrics import accuracy_score
predictions = model.predict(dtest)
accuracy = accuracy_score(y_test, predictions)
print("Accuracy: %.2f%%" % (accuracy * 100.0))

from sklearn.metrics import cohen_kappa_score
kappa = cohen_kappa_score(y_test, predictions)
print("Kappa: %.2f%%" % (kappa))

# save to JSON
model.save_model("/Users/mac/PycharmProjects/XGBoost/models/model_A.json")
```
## SHAP analysis of xgboost results
```{python}
explainer_shap = shap.TreeExplainer(final_model)
shap_values = explainer_shap(X_coords)

shap_interaction_values = shap.TreeExplainer(
    final_model).shap_interaction_values(X_coords)

X_names = ['pct_18_34','pct_white','pct_bach',
          'pct_no_car','log_pop_den','job_entropy','network_den',
          'ave_trip_dist','pct_share','x_coord','y_coord']

fig, ax = plt.subplots(figsize=(8, 6),dpi=160)

shap.summary_plot(shap_interaction_values, X_coords, max_display=15, 
                  feature_names = X_names,
                  plot_type="compact_dot")
```

### 2022-06-27 ELSA.R

@@ -1,0 +1,9 @@
+library(elsa)
+#> Loading required package: sp
+#> Loading required package: raster
+
+file <- system.file('external/dem_example.grd',package='elsa') 
+
+r <- raster(file) # reading a raster map (Dogital Elevation Model: DEM)
+
+plot(r, main='DEM: a continuous raster map')

### 2022-06-28 paper3.Rmd
hyperparameter tuning

@@
 print(y.shape, X.shape)
+X.head()
+y.head()
 # EDA_plot(y, X)
 
-ratio = 0.8
 from sklearn.model_selection import train_test_split
 # Create train/test split
-X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=1 - ratio, train_size=ratio, random_state=19, stratify=y)
+X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=0.2, random_state=2046, stratify=y)
+
 ```
+
 ## Hyperparameter Tuning
+
+The optimization process consists of 4 parts which are as follows:
+
+#### 1. Initialize domain space for range of values
+
+The domain space is the input values over which we want to search.
+
+Choose hyperparameter domain to search over
+
+-   `eta` [0,1] is analogous to learning rate in GBM conceived as the step size shrinkage.
+
+-   `n_estimators` [X/L/C]: no. of boosting iterations
+
+-   `gamma` [0,∞] specifies the minimum loss reduction required to make a split.
+
+-   `max_depth` [0,∞] refers to the maximum depth of a tree, same as GBM, which is used to control over-fitting as higher depth will allow model to learn relations very specific to a particular sample.
+
+-   `min_child_weight` [0,∞] is similar to min_child_leaf in GBM but not exactly which refers to min 'sum of weights' of observations while GBM has min 'number of observations.'
+
+-   `max_delta_step` [0,∞] is usually not needed, but it might help in logistic regression when class is extremely imbalanced.
+
+-   `subsample` [0,1] denotes the fraction of observations to be randomly samples for each tree.
+
+-   `colsample_bytree`
+
+-   `lamda` denotes L2 regularization term on weights (analogous to Ridge regression).
+
+-   `alpha` denotes L1 regularization term on weights (analogous to Lasso regression), which can be used in case of very high dimensionality so that the algorithm runs faster when implemented.
+
+-   `tree_method`
+
+-   `scale_pos_weight`
+
+-   `max_leaves`
+
 ```{python}
-%%time
+%time
 from hyperopt import fmin, tpe, hp, STATUS_OK, Trials, space_eval
+import numpy as np
 
-# Choose hyperparameter domain to search over
+# configuration spaces/structures using dictionary container type
 space = {
+  # General Parameters
+           'booster': 'gbtree',
+     # 'boosting_type': hp.choice('boosting_type', ['gbdt', 'dart']),
+     #     'verbosity':
+     #       'nthread':
 
-         'max_depth':hp.choice('max_depth', np.arange(1, 30, 1, dtype=int)),
-  'colsample_bytree':hp.quniform('colsample_bytree', 0.3, 1.01, 0.05),
-  'min_child_weight':hp.choice('min_child_weight', np.arange(1, 30, 1, dtype=int)),
-         'subsample':hp.quniform('subsample', 0.3, 1.01, 0.05),
-     'learning_rate':hp.choice('learning_rate',    np.arange(0.05, 1.01, 0.05)),
+  # Booster Parameters      
+               'eta': hp.uniform('eta', 0.01, 1.01),
              'gamma': hp.quniform('gamma', 0.1, 5, 0.05), 
-         # 'objective':'reg:squarederror',
+         'max_depth': hp.choice('max_depth', np.arange(1, 30, 1, dtype=int)),
+  'min_child_weight': hp.choice('min_child_weight', np.arange(1, 30, 1, dtype=int)),     
+         'subsample': hp.quniform('subsample', 0.3, 1.01, 0.05),
+  'colsample_bytree': hp.uniform('colsample_bytree', 0.5, 1.0),
+             'lamda': hp.uniform('lambda', 0.0, 1.0),
+             'alpha': hp.quniform('alpha', 40, 180, 1),
+  #      'tree_method': 
+  # 'scale_pos_weight':
+  #       'max_leaves':
+
+      # 'n_estimators': hp.quniform('n_estimators', 5, 35, 1),
+        # 'num_leaves': hp.quniform('num_leaves', 5, 50, 1),
+
+  # Learning Task Parameters  
          'objective': 'multi:softmax',
-       'eval_metric': 'rmse',
+       'eval_metric': 'mlogloss',
+              'seed': 2046
     }
+
+# Sample values from these configuration spaces using the routines in hyperopt.pyll.stochastic
+from hyperopt.pyll.stochastic import sample
+space = {'learning_rate': hp.loguniform('learning_rate', np.log(0.01), np.log(1))}
+print(sample(space))
+```
+
+#### 2. Define objective function
+
+The objective function can be any function which returns a real value that we want to minimize.
+In this case, we want to minimize the validation error of a machine learning model with respect to the hyperparameters.
+If the real value is accuracy, then we want to maximize it.
+Then the function should return the negative of that metric.
+
+```{python}
+from sklearn.model_selection import cross_val_score
+from hyperopt import STATUS_OK
+import lightgbm as lgb
+
+def objective_function(params):
+    clf = lgb.LGBMClassifier(**params)
+    score = cross_val_score(clf, X_train, y_train, cv=5).mean()
+    return {'loss': -score, 'status': STATUS_OK}
+
+def score(params):
+    model = XGBRegressor(**params)
+
+    model.fit(X_train, Y_train, eval_set=[(X_train, Y_train), (X_valid, Y_valid)],
+              verbose=False, early_stopping_rounds=10)
+    Y_pred = model.predict(X_valid).clip(0, 20)
+    score = sqrt(mean_squared_error(Y_valid, Y_pred))
+    print(score)
+    return {'loss': score, 'status': STATUS_OK}
+
+def optimize(trials, space):
+
+    best = fmin(score, space, algo=tpe.suggest, max_evals=1000)
+    return best
+
+trials = Trials()
+best_params = optimize(trials, space)
+
+# Return the best parameters
+space_eval(space, best_params)
+
+```
+
+From Li Ziqi
+
+```{python}
 
@@
 ```
+
+#### 3. Optimization algorithm
+
+It is the method used to construct the surrogate objective function and choose the next values to evaluate.
+
+Hyperopt has the TPE option along with random search.
+During optimization, the TPE algorithm constructs the probability model from the past results and decides the next set of hyperparameters to evaluate in the objective function by maximizing the expected improvement.
+
+```{python}
+# Step 3: Optimization algorithm (Tree Parzen Estimator, tpe)
+from hyperopt import tpe
+
+tpe_algorithm = tpe.suggest
+```
+
+#### 4. Results
+
+Results are score or value pairs that the algorithm uses to build the model.
 
@@
 ```
+
 ## SHAP analysis of xgboost results
+
 ```{python}

### 2022-06-28 paper3.Rmd
objective function for multi-classification

@@
 ```{python}
+# https://notebooks.githubusercontent.com/view/ipynb?browser=chrome&color_mode=auto&commit=daa2504c337a74778632d44834209425cb6d357f&device=unknown&enc_url=68747470733a2f2f7261772e67697468756275736572636f6e74656e742e636f6d2f4d696c6164536861686964692f46726175642d446574656374696f6e2d5847426f6f73742f646161323530346333333761373437373836333264343438333432303934323563623664333537662f5847426f6f73745f46726175645f446574656374696f6e2e6970796e62&logged_in=false&nwo=MiladShahidi%2FFraud-Detection-XGBoost&path=XGBoost_Fraud_Detection.ipynb&platform=android&repository_id=152300978&repository_type=Repository&version=98
+def objective(params, X, y, X_early_stop, y_early_stop, scorer, n_folds = 10):
+
+    pos_count = y_train.sum()
+    neg_count = len(y_train) - pos_count
+    imbalance_ratio = neg_count / pos_count
+
+    xgb_clf = XGBClassifier(**params, scale_pos_weight=imbalance_ratio,
+                            n_estimators = 2000, n_jobs = 1)
+
+    xgb_fit_params = {'early_stopping_rounds': 50,
+                      'eval_metric': ['logloss'],
+                      'eval_set': [(X_early_stop, y_early_stop)],
+                      'verbose': False
+                      }
+
+    cv_score = np.mean(cross_val_score(xgb_clf, X_train, y_train, cv = n_folds,
+                               fit_params = xgb_fit_params, n_jobs = -1,
+                               scoring = scorer))
+
+    # hypoeropt minimizes the loss, hence the minus sign behind cv_score
+    return {'loss': -cv_score, 'status': hyperopt.STATUS_OK, 'params': params}
+```
+
+```{python}
 from sklearn.model_selection import cross_val_score

### 2022-06-30 paper3.Rmd
hyper

--- 2022-06-28 paper3.Rmd
@@
 ```{python}
-%time
+#%time
 from hyperopt import fmin, tpe, hp, STATUS_OK, Trials, space_eval
@@
 # configuration spaces/structures using dictionary container type
-space = {
+param_space = {
   # General Parameters
@@
 Then the function should return the negative of that metric.
+
+```{python}
+import xgboost as xgb
+
+# https://notebooks.githubusercontent.com/view/ipynb?browser=chrome&color_mode=auto&commit=daa2504c337a74778632d44834209425cb6d357f&device=unknown&enc_url=68747470733a2f2f7261772e67697468756275736572636f6e74656e742e636f6d2f4d696c6164536861686964692f46726175642d446574656374696f6e2d5847426f6f73742f646161323530346333333761373437373836333264343438333432303934323563623664333537662f5847426f6f73745f46726175645f446574656374696f6e2e6970796e62&logged_in=false&nwo=MiladShahidi%2FFraud-Detection-XGBoost&path=XGBoost_Fraud_Detection.ipynb&platform=android&repository_id=152300978&repository_type=Repository&version=98
+def objective(params, X = X_train, y = y_train, X_early_stop = 100, y_early_stop = 100, n_folds = 10):
+
+    # pos_count = y_train.sum()
+    # neg_count = len(y_train) - pos_count
+    # imbalance_ratio = neg_count / pos_count
+
+    xgb_clf = xgb.Classifier(**params, n_estimators = 2000, n_jobs = 1)
+
+    xgb_fit_params = {'early_stopping_rounds': 50,
+                      'eval_metric': ['logloss'],
+                      'eval_set': [(X_early_stop, y_early_stop)],
+                      'verbose': True
+                      }
+
+    # cv_score = np.mean(cross_val_score(xgb_clf, X_train, y_train, cv = n_folds,
+    #                            fit_params = xgb_fit_params, n_jobs = -1,
+    #                            scoring = scorer))
+
+    cv_score = np.mean(cohen_kappa_score(xgb_clf,
+                                         X_train, 
+                                         y_train, 
+                                         cv = n_folds, 
+                                         fit_params = xgb_fit_params, 
+                                         n_jobs = -1, 
+                                         scoring = scorer)
+                       )
+
+    # hypoeropt minimizes the loss, hence the minus sign behind cv_score
+    return {'loss': cv_score, 'status': hyperopt.STATUS_OK, 'params': params}
+
+```
+
+```{python}
+# Classifier:
+def hyperparameter_tuning(space):
+    model = xgb.XGBClassifier(n_estimators = space['n_estimators'], 
+                              max_depth = int(space['max_depth']), 
+                              gamma = space['gamma'],
+                              reg_alpha = int(space['reg_alpha']),
+                              min_child_weight = space['min_child_weight'],
+                              colsample_bytree=space['colsample_bytree'])
+
+    evaluation = [( x_train, y_train), ( x_test, y_test)]
+
+    model.fit(x_train, y_train,
+            eval_set=evaluation, eval_metric="rmse",
+            early_stopping_rounds=10,verbose=False)
+
+    pred = model.predict(x_test)
+    accuracy = accuracy_score(y_test, pred>0.5)
+    print ("SCORE:", accuracy)
+    #change the metric if you like
+    return {'loss': -accuracy, 'status': STATUS_OK, 'model': model}
+
+```
+
+##### Implementing/Choosing the evaluation metric:
+
+#### 3. Optimization algorithm (Tree Parzen Estimator, tpe)
+
+It is the method used to construct the surrogate objective function and choose the next values to evaluate.
+
+Hyperopt has the TPE option along with random search.
+During optimization, the TPE algorithm constructs the probability model from the past results and decides the next set of hyperparameters to evaluate in the objective function by maximizing the expected improvement.
+
+##### Parameter Tuning: Optimizing the Cross Validated Score
+
+We now need to put these together and call Hyperopt's optimization function, fmin, to search for the optimal parameter combination.
+I will define a new function that implements this optimization and returns the parameters that acheive the highest cross validated score.
+
+The fmin function minimzes the objective function over the paramter space defined by param_space (described below).
+It takes a Trials object to store the results of all iterations, which can be used later to gain insights about the search process and results.
+For more about this (and Tree-structured Parzen Estimator, TPE, which Hyperopt uses internally for optimization) see Hyperopt documentation.
+
+```{python}
+def tune_xgb(param_space, X_train, y_train, X_early_stop = 100, y_early_stop = 100, n_iter = 100):    
+    # scorer = make_scorer(conditional_recall_score, needs_proba=True)
+
+    # hyperopt.fmin will only pass the parameter values to objective. So we need to
+    # create a partial function to bind the rest of the arguments we want to pass to objective
+    # obj = partial(objective,
+    #               # scorer = scorer,
+    #               X = X_train,
+    #               y = y_train,
+    #               X_early_stop = X_early_stop,
+    #               y_early_stop = y_early_stop)
+
+    # A trials object that will store the results of all iterations
+    trials = Trials()
+
+    fmin(fn = objective, 
+         space = param_space, 
+         algo = tpe.suggest, 
+         max_evals = n_iter, 
+         trials = trials)
+
+    # returns the values of parameters from the best trial
+    return trials.best_trial['result']['params']
+
+```
+
+```{python}
+tune_xgb(param_space, X_train, y_train, 100, 100, 100)
+```
+
+```{python}
+trials = Trials()
+best = fmin(fn=hyperparameter_tuning,
+            space=space,
+            algo=tpe.suggest,
+            max_evals=100,
+            trials=trials)
+
+print (best)
+
+```
+
+Tuning hyper-parameters of the XGBoost classifier using Bayesian Optimization
+
+```{python}
+
+if __name__ == "__main__":    
+    # Loading the data
+    data = pd.read_csv('creditcard.csv')
+    X = data.drop('Class', axis = 1)
+    y = data['Class']
+
+    # Train/test split, 80/20, random_state set for reproduibility
+    X_train, X_test, y_train, y_test = train_test_split(X, y, stratify = y, test_size = 0.2, random_state = 1)
+
+    # Further splitting the initial training set so that 10% of all data(1/8 of 80%) 
+    # can be used as the evaluation set by XGBoost for early stopping
+    X_train, X_early_stop, y_train, y_early_stop = train_test_split(X_train, y_train, stratify = y_train, test_size = 1/8, random_state = 1)
+
+    # The prior probability distribution of parameters for Bayesian optimization
+
+    param_space = {
+  # General Parameters
+           'booster': 'gbtree',
+     # 'boosting_type': hp.choice('boosting_type', ['gbdt', 'dart']),
+     #     'verbosity':
+     #       'nthread':
+
+  # Booster Parameters      
+               'eta': hp.uniform('eta', 0.01, 1.01),
+             'gamma': hp.quniform('gamma', 0.1, 5, 0.05), 
+         'max_depth': hp.choice('max_depth', np.arange(1, 30, 1, dtype=int)),
+  'min_child_weight': hp.choice('min_child_weight', np.arange(1, 30, 1, dtype=int)),     
+         'subsample': hp.quniform('subsample', 0.3, 1.01, 0.05),
+  'colsample_bytree': hp.uniform('colsample_bytree', 0.5, 1.0),
+             'lamda': hp.uniform('lambda', 0.0, 1.0),
+             'alpha': hp.quniform('alpha', 40, 180, 1),
+  #      'tree_method': 
+  # 'scale_pos_weight':
+  #       'max_leaves':
+
+      # 'n_estimators': hp.quniform('n_estimators', 5, 35, 1),
+        # 'num_leaves': hp.quniform('num_leaves', 5, 50, 1),
+
+  # Learning Task Parameters  
+         'objective': 'multi:softmax',
+       'eval_metric': 'mlogloss',
+              'seed': 2046
+    }
+
+    param_space = {
+            'learning_rate': hyperopt.hp.loguniform('learning_rate', np.log(0.01), np.log(0.2)),
+            'subsample': hyperopt.hp.uniform('subsample', 0.25, 1),
+            'colsample_bytree': hyperopt.hp.uniform('colsample_bytree', 0.7, 1.0),
+            'min_child_weight': hyperopt.hp.choice('min_child_weight', [1, 3, 5, 7]),
+            'reg_alpha': hyperopt.hp.uniform('reg_alhpa', 0, 1.0),
+            # Avoiding lambda = 0. There is a Github issue on strange behaviour with lambda = 0
+            'reg_lambda': hyperopt.hp.uniform('reg_lambda', 0.01, 1.0),
+            }
+
+      from hyperopt.pyll.stochastic import sample
+space = {'subsample': hp.uniform('subsample', 0.25, 1)}
+print(sample(space))
+
+      best_params = tune_xgb(param_space, X_train, y_train, X_early_stop, y_early_stop, n_iter = 150)
+      print('\tThe best hyper-parameters found:\n')print(*['\t\t%s = %s' % (k, str(round(v, 4))) for k, v in best_params.items()], sep='\n')
+```
+
+other references
 
@@
     return best
-```
-
-#### 3. Optimization algorithm
-
-It is the method used to construct the surrogate objective function and choose the next values to evaluate.
-
-Hyperopt has the TPE option along with random search.
-During optimization, the TPE algorithm constructs the probability model from the past results and decides the next set of hyperparameters to evaluate in the objective function by maximizing the expected improvement.
-
-```{python}
-# Step 3: Optimization algorithm (Tree Parzen Estimator, tpe)
-from hyperopt import tpe
-
-tpe_algorithm = tpe.suggest
 ```

### 2022-07-08 paper3.Rmd
objective function

@@
 
-#### 1. Initialize domain space for range of values
+#### 1. Define a search space
 
@@
 
-#### 2. Define objective function
+#### 2. Define an objective function
 
@@
 import xgboost as xgb
+from sklearn.metrics import cohen_kappa_score
+from hyperopt import fmin, tpe, hp, STATUS_OK, Trials, space_eval
 
 # https://notebooks.githubusercontent.com/view/ipynb?browser=chrome&color_mode=auto&commit=daa2504c337a74778632d44834209425cb6d357f&device=unknown&enc_url=68747470733a2f2f7261772e67697468756275736572636f6e74656e742e636f6d2f4d696c6164536861686964692f46726175642d446574656374696f6e2d5847426f6f73742f646161323530346333333761373437373836333264343438333432303934323563623664333537662f5847426f6f73745f46726175645f446574656374696f6e2e6970796e62&logged_in=false&nwo=MiladShahidi%2FFraud-Detection-XGBoost&path=XGBoost_Fraud_Detection.ipynb&platform=android&repository_id=152300978&repository_type=Repository&version=98
-def objective(params, X = X_train, y = y_train, X_early_stop = 100, y_early_stop = 100, n_folds = 10):
+def objective(params, X_train = X_train, y_train = y_train, X_test = X_test, y_test = y_test):
 
-    # pos_count = y_train.sum()
-    # neg_count = len(y_train) - pos_count
-    # imbalance_ratio = neg_count / pos_count
+    model = xgb.XGBClassifier(params,
+                              eval_metric = "rmse", 
+                              n_estimators = 2000)
 
-    xgb_clf = xgb.Classifier(**params, n_estimators = 2000, n_jobs = 1)
+    model.fit(X_train, y_train)
 
-    xgb_fit_params = {'early_stopping_rounds': 50,
-                      'eval_metric': ['logloss'],
-                      'eval_set': [(X_early_stop, y_early_stop)],
-                      'verbose': True
-                      }
+    y_pred = model.predict(X_test)
+    score = np.mean(cohen_kappa_score(y_test, y_pred))
 
-    # cv_score = np.mean(cross_val_score(xgb_clf, X_train, y_train, cv = n_folds,
-    #                            fit_params = xgb_fit_params, n_jobs = -1,
-    #                            scoring = scorer))
-
-    cv_score = np.mean(cohen_kappa_score(xgb_clf,
-                                         X_train, 
-                                         y_train, 
-                                         cv = n_folds, 
-                                         fit_params = xgb_fit_params, 
-                                         n_jobs = -1, 
-                                         scoring = scorer)
-                       )
-
-    # hypoeropt minimizes the loss, hence the minus sign behind cv_score
-    return {'loss': cv_score, 'status': hyperopt.STATUS_OK, 'params': params}
-
-```
-
-```{python}
-# Classifier:
-def hyperparameter_tuning(space):
-    model = xgb.XGBClassifier(n_estimators = space['n_estimators'], 
-                              max_depth = int(space['max_depth']), 
-                              gamma = space['gamma'],
-                              reg_alpha = int(space['reg_alpha']),
-                              min_child_weight = space['min_child_weight'],
-                              colsample_bytree=space['colsample_bytree'])
-
-    evaluation = [( x_train, y_train), ( x_test, y_test)]
-
-    model.fit(x_train, y_train,
-            eval_set=evaluation, eval_metric="rmse",
-            early_stopping_rounds=10,verbose=False)
-
-    pred = model.predict(x_test)
-    accuracy = accuracy_score(y_test, pred>0.5)
-    print ("SCORE:", accuracy)
-    #change the metric if you like
-    return {'loss': -accuracy, 'status': STATUS_OK, 'model': model}
+    return {'loss': score, 'status': STATUS_OK}
 
@@
 
-#### 3. Optimization algorithm (Tree Parzen Estimator, tpe)
+#### 3. Minimize the objective over the space
 
@@
 ```{python}
-def tune_xgb(param_space, X_train, y_train, X_early_stop = 100, y_early_stop = 100, n_iter = 100):    
-    # scorer = make_scorer(conditional_recall_score, needs_proba=True)
+from hyperopt import fmin, tpe
 
-    # hyperopt.fmin will only pass the parameter values to objective. So we need to
-    # create a partial function to bind the rest of the arguments we want to pass to objective
-    # obj = partial(objective,
-    #               # scorer = scorer,
-    #               X = X_train,
-    #               y = y_train,
-    #               X_early_stop = X_early_stop,
-    #               y_early_stop = y_early_stop)
-
-    # A trials object that will store the results of all iterations
 trials = Trials()
-
-    fmin(fn = objective, 
+best = fmin(fn = objective,
             space = param_space,
             algo = tpe.suggest, 
-         max_evals = n_iter, 
-         trials = trials)
+            max_evals = 10)
 
-    # returns the values of parameters from the best trial
-    return trials.best_trial['result']['params']
-
+print(best)
 ```

### 2022-07-08 paper3.Rmd
AUC

@@
     model = xgb.XGBClassifier(params,
-                              eval_metric = "rmse", 
+                             # eval_metric = "rmse", 
                               n_estimators = 2000)
 
-    model.fit(X_train, y_train)
+    model.fit(X_train, y_train,
+              eval_metric = 'auc')
 
@@
     return {'loss': score, 'status': STATUS_OK}
+
+```
+
+```{python}
+def objective(space):
+    clf=xgb.XGBClassifier(n_estimators =space['n_estimators'], 
+                          max_depth = int(space['max_depth']), 
+                          gamma = space['gamma'],
+                          reg_alpha =  int(space['reg_alpha']),
+                          min_child_weight=int(space['min_child_weight']),
+                          colsample_bytree=int(space['colsample_bytree']))
+
+    evaluation = [( X_train, y_train), ( X_test, y_test)]
+
+    clf.fit(X_train, y_train,
+            eval_set=evaluation, eval_metric="auc",
+            early_stopping_rounds=10,verbose=False)
+
+    pred = clf.predict(X_test)
+    accuracy = accuracy_score(y_test, pred>0.5)
+    print ("SCORE:", accuracy)
+    return {'loss': -accuracy, 'status': STATUS_OK }
 

### 2022-07-11 paper3.Rmd
update Hyperopt

--- 2022-07-08 paper3.Rmd
@@
 -   `eta` [0,1] is analogous to learning rate in GBM conceived as the step size shrinkage.
-
 -   `n_estimators` [X/L/C]: no. of boosting iterations
-
 -   `gamma` [0,∞] specifies the minimum loss reduction required to make a split.
-
 -   `max_depth` [0,∞] refers to the maximum depth of a tree, same as GBM, which is used to control over-fitting as higher depth will allow model to learn relations very specific to a particular sample.
-
 -   `min_child_weight` [0,∞] is similar to min_child_leaf in GBM but not exactly which refers to min 'sum of weights' of observations while GBM has min 'number of observations.'
-
 -   `max_delta_step` [0,∞] is usually not needed, but it might help in logistic regression when class is extremely imbalanced.
-
 -   `subsample` [0,1] denotes the fraction of observations to be randomly samples for each tree.
-
 -   `colsample_bytree`
-
 -   `lamda` denotes L2 regularization term on weights (analogous to Ridge regression).
-
 -   `alpha` denotes L1 regularization term on weights (analogous to Lasso regression), which can be used in case of very high dimensionality so that the algorithm runs faster when implemented.
-
 -   `tree_method`
-
 -   `scale_pos_weight`
-
 -   `max_leaves`
@@
 ```{python}
-#%time
+# -- hyperopt for optimising a 
 from hyperopt import fmin, tpe, hp, STATUS_OK, Trials, space_eval
+from sklearn.metrics import cohen_kappa_score
 import numpy as np
+import xgboost as xgb
 
-# configuration spaces/structures using dictionary container type
+# -- define a search space
 param_space = {
@@
   'colsample_bytree': hp.uniform('colsample_bytree', 0.5, 1.0),
-             'lamda': hp.uniform('lambda', 0.0, 1.0),
-             'alpha': hp.quniform('alpha', 40, 180, 1),
-  #      'tree_method': 
+        'reg_lambda': hp.uniform('reg_lambda', 0.0, 1.0),
+         'reg_alpha': hp.quniform('reg_alpha', 40, 180, 1),
+  #      'tree_method': 'hist','gpu_hist',  # Use GPU accelerated algorithm
   # 'scale_pos_weight':
   #       'max_leaves':
-
   #     'n_estimators': hp.quniform('n_estimators', 5, 35, 1),
   #       'num_leaves': hp.quniform('num_leaves', 5, 50, 1),
+  #        'predictor': 'cpu_predictor'
 
   # Learning Task Parameters
-         'objective': 'multi:softmax',
-       'eval_metric': 'mlogloss',
+         'num_class': 4,                   # Number of possible output classes
+         'objective': 'multi:softmax',     # Specify multiclass classification
+       'eval_metric': 'merror',
               'seed': 2046
@@
 # Sample values from these configuration spaces using the routines in hyperopt.pyll.stochastic
-from hyperopt.pyll.stochastic import sample
-space = {'learning_rate': hp.loguniform('learning_rate', np.log(0.01), np.log(1))}
-print(sample(space))
+# from hyperopt.pyll.stochastic import sample
+# space = {'learning_rate': hp.loguniform('learning_rate', np.log(0.01), np.log(1))}
+# print(sample(space))
 ```
 
-#### 2. Define an objective function
+#### 2. Define an objective/loss/cost function
 
@@
 ```{python}
-import xgboost as xgb
-from sklearn.metrics import cohen_kappa_score
-from hyperopt import fmin, tpe, hp, STATUS_OK, Trials, space_eval
-
-# https://notebooks.githubusercontent.com/view/ipynb?browser=chrome&color_mode=auto&commit=daa2504c337a74778632d44834209425cb6d357f&device=unknown&enc_url=68747470733a2f2f7261772e67697468756275736572636f6e74656e742e636f6d2f4d696c6164536861686964692f46726175642d446574656374696f6e2d5847426f6f73742f646161323530346333333761373437373836333264343438333432303934323563623664333537662f5847426f6f73745f46726175645f446574656374696f6e2e6970796e62&logged_in=false&nwo=MiladShahidi%2FFraud-Detection-XGBoost&path=XGBoost_Fraud_Detection.ipynb&platform=android&repository_id=152300978&repository_type=Repository&version=98
+# -- define an objective function
 def objective(params, X_train = X_train, y_train = y_train, X_test = X_test, y_test = y_test):
-
-    model = xgb.XGBClassifier(params,
-                              eval_metric = "rmse", 
-                              n_estimators = 2000)
-
-    model.fit(X_train, y_train)
-
+    model = xgb.XGBClassifier(**params)
+    model.fit(X_train, y_train,
+              eval_set = [(X_train, y_train), (X_test, y_test)])
     y_pred = model.predict(X_test)
     score = np.mean(cohen_kappa_score(y_test, y_pred))
-
+    # -- return more than just the objective value
     return {'loss': score, 'status': STATUS_OK}
-
 ```
@@
 
-#### 3. Minimize the objective over the space
+#### 3. Minimize the objective over the space (optimisation algorithm)
 
 It is the method used to construct the surrogate objective function and choose the next values to evaluate.
+The optimization algorithm is based on the Sequential Model-Based Global Optimization (SMBO) methodology with the variants given by the Gaussian Processes (GP), Tree of Parzen Estimators (TPE) and Adaptive Tree of Parzen Estimators (ATPE) algorithms.
 
@@
 ```{python}
-from hyperopt import fmin, tpe
-
-trials = Trials()
+# -- optimize it using algorithm of choice
+def optimize(trials, space):
     best = fmin(fn = objective,
                 space = param_space, 
-            algo = tpe.suggest, 
-            max_evals = 10)
-
-print(best)
-```
-
-```{python}
-tune_xgb(param_space, X_train, y_train, 100, 100, 100)
-```
-
-```{python}
-trials = Trials()
-best = fmin(fn=hyperparameter_tuning,
-            space=space,
-            algo=tpe.suggest,
-            max_evals=100,
-            trials=trials)
-
-print (best)
-
-```
-
-Tuning hyper-parameters of the XGBoost classifier using Bayesian Optimization
-
-```{python}
-
-if __name__ == "__main__":    
-    # Loading the data
-    data = pd.read_csv('creditcard.csv')
-    X = data.drop('Class', axis = 1)
-    y = data['Class']
-
-    # Train/test split, 80/20, random_state set for reproduibility
-    X_train, X_test, y_train, y_test = train_test_split(X, y, stratify = y, test_size = 0.2, random_state = 1)
-
-    # Further splitting the initial training set so that 10% of all data(1/8 of 80%) 
-    # can be used as the evaluation set by XGBoost for early stopping
-    X_train, X_early_stop, y_train, y_early_stop = train_test_split(X_train, y_train, stratify = y_train, test_size = 1/8, random_state = 1)
-
-    # The prior probability distribution of parameters for Bayesian optimization
-
-    param_space = {
-  # General Parameters
-           'booster': 'gbtree',
-     # 'boosting_type': hp.choice('boosting_type', ['gbdt', 'dart']),
-     #     'verbosity':
-     #       'nthread':
-
-  # Booster Parameters      
-               'eta': hp.uniform('eta', 0.01, 1.01),
-             'gamma': hp.quniform('gamma', 0.1, 5, 0.05), 
-         'max_depth': hp.choice('max_depth', np.arange(1, 30, 1, dtype=int)),
-  'min_child_weight': hp.choice('min_child_weight', np.arange(1, 30, 1, dtype=int)),     
-         'subsample': hp.quniform('subsample', 0.3, 1.01, 0.05),
-  'colsample_bytree': hp.uniform('colsample_bytree', 0.5, 1.0),
-             'lamda': hp.uniform('lambda', 0.0, 1.0),
-             'alpha': hp.quniform('alpha', 40, 180, 1),
-  #      'tree_method': 
-  # 'scale_pos_weight':
-  #       'max_leaves':
-
-      # 'n_estimators': hp.quniform('n_estimators', 5, 35, 1),
-        # 'num_leaves': hp.quniform('num_leaves', 5, 50, 1),
-
-  # Learning Task Parameters  
-         'objective': 'multi:softmax',
-       'eval_metric': 'mlogloss',
-              'seed': 2046
-    }
-
-    param_space = {
-            'learning_rate': hyperopt.hp.loguniform('learning_rate', np.log(0.01), np.log(0.2)),
-            'subsample': hyperopt.hp.uniform('subsample', 0.25, 1),
-            'colsample_bytree': hyperopt.hp.uniform('colsample_bytree', 0.7, 1.0),
-            'min_child_weight': hyperopt.hp.choice('min_child_weight', [1, 3, 5, 7]),
-            'reg_alpha': hyperopt.hp.uniform('reg_alhpa', 0, 1.0),
-            # Avoiding lambda = 0. There is a Github issue on strange behaviour with lambda = 0
-            'reg_lambda': hyperopt.hp.uniform('reg_lambda', 0.01, 1.0),
-            }
-
-      from hyperopt.pyll.stochastic import sample
-space = {'subsample': hp.uniform('subsample', 0.25, 1)}
-print(sample(space))
-
-      best_params = tune_xgb(param_space, X_train, y_train, X_early_stop, y_early_stop, n_iter = 150)
-      print('\tThe best hyper-parameters found:\n')print(*['\t\t%s = %s' % (k, str(round(v, 4))) for k, v in best_params.items()], sep='\n')
-```
-
-other references
-
-```{python}
-from sklearn.model_selection import cross_val_score
-from hyperopt import STATUS_OK
-import lightgbm as lgb
-
-def objective_function(params):
-    clf = lgb.LGBMClassifier(**params)
-    score = cross_val_score(clf, X_train, y_train, cv=5).mean()
-    return {'loss': -score, 'status': STATUS_OK}
-
-def score(params):
-    model = XGBRegressor(**params)
-
-    model.fit(X_train, Y_train, eval_set=[(X_train, Y_train), (X_valid, Y_valid)],
-              verbose=False, early_stopping_rounds=10)
-    Y_pred = model.predict(X_valid).clip(0, 20)
-    score = sqrt(mean_squared_error(Y_valid, Y_pred))
-    print(score)
-    return {'loss': score, 'status': STATUS_OK}
-
-def optimize(trials, space):
-
-    best = fmin(score, space, algo=tpe.suggest, max_evals=1000)
+                algo = tpe.suggest,    # the optimisation algorithm
+                max_evals = 10,        # the number of iteration
+                trials = trials)       # the trials database for fmin
     return best
 
+# -- create a Trials database to store experiment results
 trials = Trials()
-best_params = optimize(trials, space)
+best_params = optimize(trials, param_space)
 
 # Return the best parameters
-space_eval(space, best_params)
+space_eval(param_space, best_params)
 
+# -- print out what we got!
+# for trial in trials.trials:
+#     print(trial)
 ```
-
-From Li Ziqi
-
-```{python}
-
-def score(params, n_folds=5):
-
-    #Cross-validation
-    d_train = xgboost.DMatrix(X_coords,y)
-
-    cv_results = xgboost.cv(params, d_train, nfold = n_folds, num_boost_round=500,
-                        early_stopping_rounds = 10, metrics = 'rmse', seed = 0)
-
-    loss = min(cv_results['test-rmse-mean'])
-
-    return loss
-
-def optimize(trials, space):
-
-    best = fmin(score, space, algo=tpe.suggest, max_evals=2000,
-                rstate=np.random.RandomState(333))#Add seed to fmin function
-    return best
-```
-
 #### 4. Results
@@
 
-# Leave most parameters as default
-param = {'objective': 'multi:softmax',  # Specify multiclass classification
-         'num_class': 4,  # Number of possible output classes
-         # 'tree_method': 'gpu_hist',  # Use GPU accelerated algorithm
-         'eta': 0.451,
-         'max_depth': 12,
-         'gamma': 0,
-         'subsample': 0.758,
-         'colsample_bytree': 1,
-         'min_child_weight': 0.2,
-         'tree_method': 'hist',
-         'predictor': 'cpu_predictor'
-         }
-
 # Convert input data from numpy to XGBoost format
-dtrain = xgb.DMatrix(X_train, label=y_train)
-dtest = xgb.DMatrix(X_test, label=y_test)
+dtrain = xgb.DMatrix(X_train, label = y_train)
+ dtest = xgb.DMatrix(X_test, label = y_test)
 
@@
 # print("GPU Training Time: %s seconds" % (str(time.time() - start)))
-model = xgb.train(param, dtrain, num_round, evals=[(dtest, 'test')])
+
+best_params['num_class'] = 4
+best_params['objective'] = 'multi:softmax'
+best_params['eval_metric'] = 'merror'
+
+model = xgb.train(best_params,
+                  dtrain,
+                  num_round = 1000)
+                  #evals = [(dtrain, 'train'),(dtest, 'test')])
+
 print(time.time() - start)
+
+y_pred = model.predict(dtest)
 
@@
 from sklearn.metrics import accuracy_score
-predictions = model.predict(dtest)
-accuracy = accuracy_score(y_test, predictions)
+
+accuracy = accuracy_score(y_test, y_pred)
 print("Accuracy: %.2f%%" % (accuracy * 100.0))
@@
 from sklearn.metrics import cohen_kappa_score
-kappa = cohen_kappa_score(y_test, predictions)
+kappa = cohen_kappa_score(y_test, y_pred)
 print("Kappa: %.2f%%" % (kappa))
@@
 # save to JSON
-model.save_model("/Users/mac/PycharmProjects/XGBoost/models/model_A.json")
+# model.save_model("/Users/mac/PycharmProjects/XGBoost/models/model_A.json")
 ```

### 2022-07-22 paper3.Rmd
Bayesian ranking

@@
 
-## Download the wildness measure and its components
-#"https://github.com/yiminchangchien/scenicness/blob/2a3f9b996c11a7c480867524664401dbea9aea2b/Wildness.RData" %>% url() %>%
-
-load("C:\\Users\\00631\\Downloads\\James\\Wildness.RData")
-
 ## Download Scenic-Or-Not data
-sc <- read_tsv("http://scenicornot.datasciencelab.co.uk/votes.tsv",
+sc1 <- read_tsv("http://scenicornot.datasciencelab.co.uk/votes.tsv",
                col_types = cols("ID" = col_number(),
@@
   mutate(Mean = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
+         Median = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
          Variance = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% var(),
          Entropy = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
-         Number = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% na.omit() %>% length())%>%
+         Number = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% length(),
+           `10` = Votes %>% strsplit(",") %>% unlist() %>% str_count("10") %>% sum(),
+            `9` = Votes %>% strsplit(",") %>% unlist() %>% str_count("9") %>% sum(),
+            `8` = Votes %>% strsplit(",") %>% unlist() %>% str_count("8") %>% sum(),
+            `7` = Votes %>% strsplit(",") %>% unlist() %>% str_count("7") %>% sum(),
+            `6` = Votes %>% strsplit(",") %>% unlist() %>% str_count("6") %>% sum(),
+            `5` = Votes %>% strsplit(",") %>% unlist() %>% str_count("5") %>% sum(),
+            `4` = Votes %>% strsplit(",") %>% unlist() %>% str_count("4") %>% sum(),
+            `3` = Votes %>% strsplit(",") %>% unlist() %>% str_count("3") %>% sum(),
+            `2` = Votes %>% strsplit(",") %>% unlist() %>% str_count("2") %>% sum(),
+            `1` = Votes %>% strsplit(",") %>% unlist() %>% str_count("1") %>% sum())%>%
+  mutate(`1` = `1`-`10`) %>%
   st_as_sf(coords = c("Lon","Lat"), crs = 4326) %>%
   st_transform(crs = 27700)
+```
+
+```{python}
+import math
+from scipy import stats
+
+def ci_lower_bound(positive_ratings, total_ratings, confidence):
+    """
+    Following http://www.evanmiller.org/how-not-to-sort-by-average-rating.html#changes
+    :return: the lower bound of the confidence interval
+    """
+    n =  total_ratings
+    z = stats.norm.ppf(1 - (1 - confidence) / 2)
+    p_hat = 1.0 * positive_ratings / n
+    result = (
+        (
+             p_hat + z * z / (2 * n)
+             - z * math.sqrt(
+                 (p_hat * (1 - p_hat) + z * z / (4 * n)) / n
+             )
+        ) / (1 + z * z / n)
+    )
+    return result
+
 ```
@@
   rowwise() %>%
-  mutate(Scenicness.median = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
-         Scenicness.mean = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
-         Scenicness.entropy = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
-         Scenicness.number = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% na.omit() %>% length(),
+  mutate(Scenicness.mean_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
+         Scenicness.median_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
+         Scenicness.entropy_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
+         Scenicness.number_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% na.omit() %>% length()) %>%
+  ungroup()
+
+LANDMAP <- sc %>%
+  st_join(LANDMAP, .) %>%
+  aggregate(Median~UID, data = ., paste0, collapse = ",") %>%
+  merge(LANDMAP, ., by = "UID", all = T) %>%
+  as_tibble() %>%
+  st_as_sf() %>%
+  rowwise() %>%
+  mutate(Scenicness.mean_of_median = Median %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
+         Scenicness.median_of_median = Median %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median()) %>%
+  ungroup()
+
+LANDMAP <- sc %>%
+  st_join(LANDMAP, .) %>%
+  aggregate(Mean~UID, data = ., paste0, collapse = ",") %>%
+  merge(LANDMAP, ., by = "UID", all = T) %>%
+  as_tibble() %>%
+  st_as_sf() %>%
+  rowwise() %>%
+  mutate(Scenicness.mean_of_mean = Mean %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
+         Scenicness.median_of_mean = Mean %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
          Consultant = ifelse(str_detect(VS_1a, "Bronwen Thomas"), "B", "A")) %>%
-  dplyr::select(UID, CLS_1, CLS_2, CLS_3, CLS_4, ScenicQuality, Consultant, Votes, Scenicness.median, Scenicness.mean, Scenicness.entropy, Scenicness.number) %>%
+  dplyr::select(UID, CLS_1, CLS_2, CLS_3, CLS_4, ScenicQuality, Consultant, 
+                Votes, Scenicness.mean_all, Scenicness.median_all, Scenicness.entropy_all, Scenicness.number_all,
+                Scenicness.mean_of_median, Scenicness.median_of_median,
+                Scenicness.mean_of_mean, Scenicness.median_of_mean) %>%
   ungroup()
@@
   dplyr::select(Wildness.mean = mean, Wildness.median = median, Wildness.iqr, Wildness.cv = coefficient_of_variation, Wildness.variance = variance) %>%
-  bind_cols(LANDMAP, .) %>%
+  bind_cols(LANDMAP[, 1:13], .) %>%
   st_transform(crs = 27700)
 
-landmap <- st_cast(LANDMAP, "POLYGON")
+landmap <- st_cast(LANDMAP, "POLYGON") %>%
+  dplyr::mutate(x = st_centroid(landmap$geometry) %>% st_coordinates() %>% .[, 1],
+                y = st_centroid(landmap$geometry) %>% st_coordinates() %>% .[, 2])
+summary(landmap)
 ```
@@
 ```{r}
+Sys.setenv("RETICULATE_PYTHON" = ".\\.venv\\Scripts\\python.exe")
+# Sys.getenv()
+Sys.which("Python")
+
+# py_discover_config()
+
 library(reticulate)
+# -- creates an interactive Python console, REPL (Read, Eval, Print Loop), within R.
+repl_python()
 
+# reticulate::py_config() # why numpy_version couldn't be found?
+# use_python("Z:\\R\\R projects\\LCA-GB\\my_env_Python_3.9.13\\Scripts\\python.exe")
+# 
+# # use_virtualenv("Z:\\R\\R projects\\LCA-GB\\my_env_Python_3.9.13", required = TRUE)
+# 
+# install.packages("usethis")
+# usethis::edit_r_environ()
+# usethis::edit_r_profile()
 ```
@@
 ```{python}
-
 import geopandas as gpd
@@
 
-# landmap = geopandas.read_file("/Users/mac/R/LCA-GB/Data/Input/LANDMAP.geojson")
 landmap = r.landmap
@@
 print(landmap)
-print(landmap.describe)
+landmap.describe()
 print(landmap['ScenicQuality'].value_counts())
@@
 # split train/test set
-X = landmap.loc[:, ['Scenicness.mean']]
-y = landmap.loc[:, ['ScenicQuality']]
+# X = landmap_B.loc[:, ['Scenicness.median', 'Scenicness.entropy', 
+#                       'Wildness.median', 'Wildness.variance',
+#                       'x', 'y']]
+
+X = landmap_B.loc[:, ['Scenicness.mean', 'Wildness.mean']]
+y = landmap_B.loc[:, ['ScenicQuality']]
 
@@
 
-scale_mapper = {'Low': int(0), 'Moderate': int(1), 'High': int(2), 'Outstanding': int(3)} 
+scale_mapper = {'Low': int(0), 
+                'Moderate': int(1), 
+                'High': int(2), 
+                'Outstanding': int(3)}
+
 y[['ScenicQuality']] = y[['ScenicQuality']].replace(scale_mapper).astype('int')
@@
 
+# -- Create train/test split
 from sklearn.model_selection import train_test_split
-# Create train/test split
-X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=0.2, random_state=2046, stratify=y)
-
+X_train, X_test, y_train, y_test = train_test_split(X, y, 
+                                                    test_size = 0.2, 
+                                                    random_state = 2046, 
+                                                    stratify = y)
 ```
-
 ## Hyperparameter Tuning
@@
 The domain space is the input values over which we want to search.
-
 Choose hyperparameter domain to search over
@@
 ```{python}
-# -- hyperopt for optimising a 
+# -- hyperopt for optimising a XGBoost model
 from hyperopt import fmin, tpe, hp, STATUS_OK, Trials, space_eval
@@
   # Booster Parameters
+   'num_boost_round': hp.quniform('num_boost_round', 100, 1500, 0.001), 
                'eta': hp.uniform('eta', 0.01, 1.01),
-             'gamma': hp.quniform('gamma', 0.1, 5, 0.05), 
-         'max_depth': hp.choice('max_depth', np.arange(1, 30, 1, dtype=int)),
-  'min_child_weight': hp.choice('min_child_weight', np.arange(1, 30, 1, dtype=int)),     
-         'subsample': hp.quniform('subsample', 0.3, 1.01, 0.05),
-  'colsample_bytree': hp.uniform('colsample_bytree', 0.5, 1.0),
-        'reg_lambda': hp.uniform('reg_lambda', 0.0, 1.0),
-         'reg_alpha': hp.quniform('reg_alpha', 40, 180, 1),
+             'gamma': 0, 
+         'max_depth': hp.choice('max_depth', np.arange(1, 10, 1, dtype=int)),
+  'min_child_weight': hp.quniform('min_child_weight', 0, 1, 0.1),     
+         'subsample': hp.quniform('subsample', 0.3, 1, 0.001),
+  'colsample_bytree': 1, #hp.uniform('colsample_bytree', 0.5, 1.01),
+ #       'reg_lambda': hp.uniform('reg_lambda', 0.0, 1.0),
+ #        'reg_alpha': hp.quniform('reg_alpha', 40, 180, 1),
  #      'tree_method': 'hist','gpu_hist',  # Use GPU accelerated algorithm
@@
  #       'num_leaves': hp.quniform('num_leaves', 5, 50, 1),
-  #        'predictor': 'cpu_predictor'
+ #        'predictor': 'cpu_predictor',
 
@@
          'num_class': 4,                   # Number of possible output classes
-         'objective': 'multi:softmax',     # Specify multiclass classification
+         'objective': 'multi:softmax',     # Specify multi-class classification
        'eval_metric': 'merror',
-              'seed': 2046
+              'seed': 2022
     }
@@
 # Sample values from these configuration spaces using the routines in hyperopt.pyll.stochastic
-# from hyperopt.pyll.stochastic import sample
-# space = {'learning_rate': hp.loguniform('learning_rate', np.log(0.01), np.log(1))}
-# print(sample(space))
+from hyperopt.pyll.stochastic import sample
+space = {'eta': hp.uniform('eta', 0.01, 1.01)}
+print(sample(space))
 ```
-
 #### 2. Define an objective/loss/cost function
@@
                 algo = tpe.suggest,    # the optimisation algorithm
-                max_evals = 10,        # the number of iteration
+                max_evals = 30,        # the number of iteration
                 trials = trials)       # the trials database for fmin
@@
 space_eval(param_space, best_params)
-
-# -- print out what we got!
-# for trial in trials.trials:
-#     print(trial)
 ```
@@
 import xgboost as xgb
-# Specify sufficient boosting iterations to reach a minimum
-num_round = 1000 #209.91
+## Specify sufficient boosting iterations to reach a minimum
+## num_round = 1000
 
-# Convert input data from numpy to XGBoost format
+# -- convert input data from numpy to XGBoost format
 dtrain = xgb.DMatrix(X_train, label = y_train)
@@
 
+# -- try previous combinations of hyperparameter set
+# Consultant A
+# param_space = {
+#            'booster': 'gbtree',
+#    'num_boost_round': 209.910,
+#                'eta': 0.451,
+#              'gamma': 0,
+#          'max_depth': 12,
+#   'min_child_weight': 0.2,
+#          'subsample': 0.758,
+#   'colsample_bytree': 1,
+#          'num_class': 4,
+#          'objective': 'multi:softmax',
+#        'eval_metric': 'merror',
+#               'seed': 2046
+#     }
+
+# Consultant B
+params = {
+          'booster': 'gbtree',
+        #'num_round': 1433.996,
+              'eta': 0.892,
+            'gamma': 0,
+        'max_depth': 4,
+ 'min_child_weight': 0.6,
+        'subsample': 0.500,
+ 'colsample_bytree': 1,
+        'num_class': 4,
+        'objective': 'multi:softmax',
+      'eval_metric': 'merror',
+             'seed': 1997
+  }
+
+# -- train model
 import time
-gpu_res = {}  # Store accuracy result
 start = time.time()
 
-# Train model
-# model = xgb.train(param, dtrain, num_round, evals=[(dtest, 'test')], evals_result=gpu_res)
-# print("GPU Training Time: %s seconds" % (str(time.time() - start)))
+model = xgb.train(space_eval(param_space, best_params), 
+                  dtrain, 
+                  num_boost_round = 500,
+                  evals = [(dtrain, 'train'), (dtest, 'test')],
+                  early_stopping_rounds = 250)
 
-best_params['num_class'] = 4
-best_params['objective'] = 'multi:softmax'
-best_params['eval_metric'] = 'merror'
+# model = xgb.train(params, 
+#                   dtrain, 
+#                   num_boost_round = 500, 
+#                   evals = [(dtrain, 'train'), (dtest, 'test')],
+#                   early_stopping_rounds = 250)
 
-model = xgb.train(best_params,
-                  dtrain,
-                  num_round = 1000)
-                  #evals = [(dtrain, 'train'),(dtest, 'test')])
+# model = xgb.cv(space_eval(param_space, best_params), dtrain, num_boost_round = 2000, nfold = 5, stratified = True)
+# model.head()
 
-print(time.time() - start)
+print("XGBoost model training time: %s seconds" % (str(time.time() - start)))
 
@@
 
-# Evaluate model using test data
+# -- evaluate model using test data
 from sklearn.metrics import accuracy_score
@@
 from sklearn.metrics import cohen_kappa_score
-kappa = cohen_kappa_score(y_test, y_pred)
-print("Kappa: %.2f%%" % (kappa))
+kappa = cohen_kappa_score(y_pred, y_test)
+print("Kappa: %.2f" % (kappa))
 
-# save to JSON
-# model.save_model("/Users/mac/PycharmProjects/XGBoost/models/model_A.json")
+# -- save to JSON
+model.save_model("Z:\\R\\R projects\\LCA-GB\\models\\model_B.json")
 ```
@@
 ```{python}
-explainer_shap = shap.TreeExplainer(final_model)
-shap_values = explainer_shap(X_coords)
+import shap
+import numpy as np
+import matplotlib.pyplot as plt
+from sklearn.preprocessing import StandardScaler
+from sklearn.metrics import plot_confusion_matrix
 
-shap_interaction_values = shap.TreeExplainer(
-    final_model).shap_interaction_values(X_coords)
+# -- load the saved model
+model = xgb.XGBRegressor()
+model.load_model("Z:\\R\\R projects\\LCA-GB\\models\\model_B.json")
 
-X_names = ['pct_18_34','pct_white','pct_bach',
-          'pct_no_car','log_pop_den','job_entropy','network_den',
-          'ave_trip_dist','pct_share','x_coord','y_coord']
+# -- view the confusion matrix
+class_names = ['Low', 'Moderate', 'High', 'Outstanding']
+disp = plot_confusion_matrix(model, X_test, y_test, 
+                             display_labels = class_names, 
+                             cmap = plt.cm.Blues, 
+                             xticks_rotation = 'vertical')
 
-fig, ax = plt.subplots(figsize=(8, 6),dpi=160)
+# -- view the feature importance
+importances = model.feature_importances_
+indices = np.argsort(importances)
+features = X.columns
 
-shap.summary_plot(shap_interaction_values, X_coords, max_display=15, 
+fig, ax = plt.subplots()
+
+plt.title('Feature Importances')
+plt.barh(range(len(indices)), importances[indices], color='g', align='center')
+plt.yticks(range(len(indices)), [features[i] for i in indices])
+plt.xlabel('Relative Importance')
+plt.show()
+
+fig.set_size_inches(18.5, 10.5)
+fig.savefig("Z:\\R\\R projects\\LCA-GB\\models\\pyplot.png")
+plt.clf()
+
+# -- play shap
+explainer = shap.TreeExplainer(model)
+shap_values = explainer.shap_values(X_test, approximate = True)
+shap.summary_plot(shap_values[0], X_test.values, feature_names = X_test.columns)
+fig.set_size_inches(18.5, 20.5)
+fig.savefig("Z:\\R\\R projects\\LCA-GB\\models\\shap.png")
+
+# -- first trial
+shap.initjs()
+shap_values = shap.TreeExplainer(model).shap_values(X_test)
+shap.summary_plot(shap_values, X_test)
+
+# -- second trial
+scaler = StandardScaler()
+scaled_data = scaler.fit_transform(X_test.iloc[:,:-1])
+shap_values = explainer.shap_values(scaled_data.iloc[0,:-1].values.reshape(1,-1))
+shap.force_plot(explainer.expected_value[0], shap_values[0], link='logit')  # repeat changing 0 for i in range(0, 5)
+
+# -- third trial
+shap.initjs()
+class_names = ['Low', 'Moderate', ' High', 'Outstanding']
+shap.summary_plot(shap_values, X_test, plot_type="bar", class_names= class_names, feature_names = ['Scenicness', 'Wildness'])
+
+# -- plot the global importance of each feature
+shap.plots.bar(explainer)
+
+# visualize the first prediction's explanation
+shap.force_plot(explainer.expected_value[0], shap_values[0])
+
+# fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
+# shap.summary_plot(shap_values)
+# # or 
+# shap.plots.beeswarm(shap_values)
+shap.summary_plot(shap_values[2], X_test)
+
+# fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
+shap.summary_plot(shap_values, X_test, plot_type="bar")
+
+shap_interaction_values = shap.TreeExplainer(model).shap_interaction_values(X_train)
+
+X_names = ['Scenicness.median','x_coord','y_coord']
+
+shap.summary_plot(shap_interaction_values, X_train, max_display=15, 
                   feature_names = X_names,
-                  plot_type="compact_dot")
+                  plot_type = 'dot')
+                  # plot_type="compact_dot")
+
+import shap  # package used to calculate Shap values
+
+# Create object that can calculate shap values
+explainer = shap.TreeExplainer(my_model)
+
+# calculate shap values. This is what we will plot.
+# Calculate shap_values for all of val_X rather than a single row, to have more data for plot.
+shap_values = explainer.shap_values(val_X)
+
+# Make plot. Index of [1] is explained in text below.
+shap.summary_plot(shap_values[1], val_X)                  
 ```

### 2022-07-25 paper3.Rmd
Bayesian Ranking

@@
 ```
+## Bayesian XXX by Evan Miller
+Where zα/2 is the 1−α/2 quantile of a normal distribution.
+
+```{r Bayesian Ranking}
+"
+Following https://www.evanmiller.org/ranking-items-with-star-ratings.html
+return: the lower bound of the confidence interval
+"
+BayesRank <- function(votes, K, α = .05){
+  N = votes %>% strsplit(",") %>% unlist() %>% length() # total_ratings
+  s_k = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10) %>% array(dim = c(1, 10))
+  n_k = votes %>%
+    strsplit(",") %>% 
+    unlist() %>%
+    as.numeric() %>%
+    tabulate(nbins = 10) %>%
+    array(dim = c(10, 1))
+  # 1-α: confidence level (commonly 90%, 95% or 99%)
+  # confidence level 90%: α = 0.1  -> z_α/2 = z_0.05 = 1.645
+  # confidence level 95%: α = 0.05 -> z_α/2 = z_0.025= 1.96
+  # confidence level 99%: α = 0.01 -> z_α/2 = z_0.005= 2.576
+  # z is the standard normal random variable. z_α/2 is the α/2 quantile of its distribution which means     critical value.
+  z = qnorm(1-α/2)
+  S = s_k %*% (n_k + 1)/(N + K) - 
+  z*sqrt((s_k^2 %*% (n_k + 1)/(N + K) - (s_k %*% (n_k + 1)/(N + K))^2)/(N + K + 1))
+  S = as.numeric(S)
+  return(S)
+}
+```
 
@@
 ## Download Scenic-Or-Not data
-sc1 <- read_tsv("http://scenicornot.datasciencelab.co.uk/votes.tsv",
+sc <- read_tsv("http://scenicornot.datasciencelab.co.uk/votes.tsv",
                col_types = cols("ID" = col_number(),
@@
          Entropy = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
-         Number = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% length(),
-           `10` = Votes %>% strsplit(",") %>% unlist() %>% str_count("10") %>% sum(),
-            `9` = Votes %>% strsplit(",") %>% unlist() %>% str_count("9") %>% sum(),
-            `8` = Votes %>% strsplit(",") %>% unlist() %>% str_count("8") %>% sum(),
-            `7` = Votes %>% strsplit(",") %>% unlist() %>% str_count("7") %>% sum(),
-            `6` = Votes %>% strsplit(",") %>% unlist() %>% str_count("6") %>% sum(),
-            `5` = Votes %>% strsplit(",") %>% unlist() %>% str_count("5") %>% sum(),
-            `4` = Votes %>% strsplit(",") %>% unlist() %>% str_count("4") %>% sum(),
-            `3` = Votes %>% strsplit(",") %>% unlist() %>% str_count("3") %>% sum(),
-            `2` = Votes %>% strsplit(",") %>% unlist() %>% str_count("2") %>% sum(),
-            `1` = Votes %>% strsplit(",") %>% unlist() %>% str_count("1") %>% sum())%>%
-  mutate(`1` = `1`-`10`) %>%
+         BayeScore = Votes %>% BayesRank(K = 10, α = .05))%>%
   st_as_sf(coords = c("Lon","Lat"), crs = 4326) %>%
   st_transform(crs = 27700)
-```
-
-```{python}
-import math
-from scipy import stats
-
-def ci_lower_bound(positive_ratings, total_ratings, confidence):
-    """
-    Following http://www.evanmiller.org/how-not-to-sort-by-average-rating.html#changes
-    :return: the lower bound of the confidence interval
-    """
-    n =  total_ratings
-    z = stats.norm.ppf(1 - (1 - confidence) / 2)
-    p_hat = 1.0 * positive_ratings / n
-    result = (
-        (
-             p_hat + z * z / (2 * n)
-             - z * math.sqrt(
-                 (p_hat * (1 - p_hat) + z * z / (4 * n)) / n
-             )
-        ) / (1 + z * z / n)
-    )
-    return result
-
 ```
@@
   rowwise() %>%
-  mutate(Scenicness.mean_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
-         Scenicness.median_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
-         Scenicness.entropy_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
-         Scenicness.number_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% na.omit() %>% length()) %>%
+  mutate(Scenicness.mean_of_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
+         Scenicness.median_of_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
+         Scenicness.entropy_of_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
+         Scenicness.number_of_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% na.omit() %>% length()) %>%
   ungroup()
@@
   mutate(Scenicness.mean_of_mean = Mean %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
-         Scenicness.median_of_mean = Mean %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
-         Consultant = ifelse(str_detect(VS_1a, "Bronwen Thomas"), "B", "A")) %>%
-  dplyr::select(UID, CLS_1, CLS_2, CLS_3, CLS_4, ScenicQuality, Consultant, 
-                Votes, Scenicness.mean_all, Scenicness.median_all, Scenicness.entropy_all, Scenicness.number_all,
-                Scenicness.mean_of_median, Scenicness.median_of_median,
-                Scenicness.mean_of_mean, Scenicness.median_of_mean) %>%
+         Scenicness.median_of_mean = Mean %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median()) %>%
   ungroup()
 
-plot(LANDMAP["ScenicQuality"], border = 'NA', key.pos = 1)  
+LANDMAP <- sc %>%
+  st_join(LANDMAP, .) %>%
+  aggregate(BayeScore~UID, data = ., paste0, collapse = ",") %>%
+  merge(LANDMAP, ., by = "UID", all = T) %>%
+  as_tibble() %>%
+  st_as_sf() %>%
+  rowwise() %>%
+  mutate(Scenicness.mean_of_bayescore = BayeScore %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
+         Scenicness.median_of_bayescore = BayeScore %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
+         Consultant = ifelse(str_detect(VS_1a, "Bronwen Thomas"), "B", "A")) %>%
+  ungroup()
+
+plot(LANDMAP["ScenicQuality"], border = 'NA', col = RColorBrewer::brewer.pal(4, "YlOrBr"), key.pos = 1)  
 ```
@@
 LANDMAP <- st_transform(LANDMAP, crs = st_crs(Wildness)) %>%
-  exact_extract(Wildness, ., fun = c('mean', 'median', 'coefficient_of_variation', 'variance', 'quantile'), quantiles = c(0.25, 0.75)) %>%
+  exact_extract(Wildness, ., 
+                fun = c('mean', 'median', 'coefficient_of_variation', 'variance', 'quantile'), 
+                quantiles = c(0.25, 0.75)) %>%
   mutate(Wildness.iqr = q75 - q25) %>%
-  dplyr::select(Wildness.mean = mean, Wildness.median = median, Wildness.iqr, Wildness.cv = coefficient_of_variation, Wildness.variance = variance) %>%
-  bind_cols(LANDMAP[, 1:13], .) %>%
+  bind_cols(LANDMAP, .) %>%
+  dplyr::select(UID, CLS_1, CLS_2, CLS_3, CLS_4, ScenicQuality, Consultant,
+                Votes, Scenicness.mean_of_all, Scenicness.median_of_all,
+                Scenicness.entropy_of_all, Scenicness.number_of_all,
+                Scenicness.mean_of_median, Scenicness.median_of_median,
+                Scenicness.mean_of_mean, Scenicness.median_of_mean,
+                Scenicness.mean_of_bayescore, Scenicness.median_of_bayescore,
+                Wildness.mean = mean, Wildness.median = median, Wildness.iqr, 
+                Wildness.cv = coefficient_of_variation, Wildness.variance = variance) %>%
   st_transform(crs = 27700)
@@
 # split train/test set
-# X = landmap_B.loc[:, ['Scenicness.median', 'Scenicness.entropy', 
-#                       'Wildness.median', 'Wildness.variance',
-#                       'x', 'y']]
-
-X = landmap_B.loc[:, ['Scenicness.mean', 'Wildness.mean']]
+# X = landmap_B.loc[:, ['Scenicness.mean', 'Wildness.mean']]
+X = landmap_B.loc[:, ['Scenicness.median', 'x', 'y']]
 y = landmap_B.loc[:, ['ScenicQuality']]

### 2022-07-29 paper3.Rmd
shap summary plot

@@
 ```
+
 ## Bayesian XXX by Evan Miller
+
 Where zα/2 is the 1−α/2 quantile of a normal distribution.
+- Evan Miller shows a Bayesian approach to ranking 5-star ratings: - <https://stackoverflow.com/questions/1411199/what-is-a-better-way-to-sort-by-a-5-star-rating/40958702#40958702>
 
@@
 # X = landmap_B.loc[:, ['Scenicness.mean', 'Wildness.mean']]
-X = landmap_B.loc[:, ['Scenicness.median', 'x', 'y']]
+X = landmap_B.loc[:, ['Scenicness.median_of_bayescore', 'Wildness.median', 'x', 'y']]
 y = landmap_B.loc[:, ['ScenicQuality']]
@@
 ```
+
 ## Hyperparameter Tuning
@@
 
-The domain space is the input values over which we want to search.
-Choose hyperparameter domain to search over
+The domain space is the input values over which we want to search. Choose hyperparameter domain to search over
 
@@
 ```
+
 #### 2. Define an objective/loss/cost function
@@
 ```
+
 #### 4. Results
@@
 
-## Fit xgboost model
+## Fit/Train the xgboost model
 
@@
           'booster': 'gbtree',
-        #'num_round': 1433.996,
+        'num_round': 1434, #1433.996,
               'eta': 0.892,
@@
       'eval_metric': 'merror',
-             'seed': 1997
+             'seed': 1995
   }
@@
 
-model = xgb.train(space_eval(param_space, best_params), 
-                  dtrain, 
-                  num_boost_round = 500,
-                  evals = [(dtrain, 'train'), (dtest, 'test')],
-                  early_stopping_rounds = 250)
-
-# model = xgb.train(params, 
+# model = xgb.train(space_eval(param_space, best_params),
 #                   dtrain,
@@
 #                   early_stopping_rounds = 250)
+
+model = xgb.train(params,
+                  dtrain,
+                  num_boost_round = 500,
+                  evals = [(dtrain, 'train'), (dtest, 'test')],
+                  verbose_eval = 100,
+                  early_stopping_rounds = 251)
 
@@
 
-```{python}
+```{python, engine.path = '.venv/Scripts/python'}
 import shap
@@
 import matplotlib.pyplot as plt
+import matplotlib.pylab as pl
+import xgboost as xgb
 from sklearn.preprocessing import StandardScaler
@@
 
+# print the JS visualization code to the notebook
+shap.initjs()
+
 # -- load the saved model
-model = xgb.XGBRegressor()
+model = xgb.XGBClassifier()
 model.load_model("Z:\\R\\R projects\\LCA-GB\\models\\model_B.json")
+```
 
+## Confusion matrix
+```{python}
 # -- view the confusion matrix
 class_names = ['Low', 'Moderate', 'High', 'Outstanding']
-disp = plot_confusion_matrix(model, X_test, y_test, 
+disp = ConfusionMatrixDisplay.from_predictions(model, X_test, y_test, 
                              display_labels = class_names, 
@@
                              xticks_rotation = 'vertical')
+```
 
+```{python}
+# -- play shap
+explainer = shap.TreeExplainer(model)
+shap_values = explainer.shap_values(X_train, approximate=True)
+
+shap.plots.beeswarm(shap_values, show=False, color_bar=False)
+plt.colorbar()
+plt.show()
+
+# shap.summary_plot(shap_values[1], 
+#                   X_train.values, 
+#                   feature_names = X_train.columns,
+#                   plot_size = (10, 4))
+#                   
+# plt.gcf().axes[-1].set_aspect('auto')
+# plt.tight_layout()
+# As mentioned, smaller "box_aspect" value to make colorbar thicker
+# plt.gcf().axes[-1].set_box_aspect(200)
+
+```
+
+```{python}
 # -- view the feature importance
@@
 plt.clf()
+```
 
-# -- play shap
-explainer = shap.TreeExplainer(model)
-shap_values = explainer.shap_values(X_test, approximate = True)
-shap.summary_plot(shap_values[0], X_test.values, feature_names = X_test.columns)
-fig.set_size_inches(18.5, 20.5)
-fig.savefig("Z:\\R\\R projects\\LCA-GB\\models\\shap.png")
+```{python}
+# get just the explanations for the positive class
+# shap_values = shap_values[...,1]
 
-# -- first trial
+# shap.plots.bar(shap_values[...,1])
+```
+
+```{python}
 shap.initjs()
-shap_values = shap.TreeExplainer(model).shap_values(X_test)
-shap.summary_plot(shap_values, X_test)
+# shap_values[1] what does the index mean?
+shap.summary_plot(shap_values[1], X_train.values, feature_names=X_train.columns)
+# shap.summary_plot(shap_values, X_test, max_display = X_test.shape[1])
 
-# -- second trial
-scaler = StandardScaler()
-scaled_data = scaler.fit_transform(X_test.iloc[:,:-1])
-shap_values = explainer.shap_values(scaled_data.iloc[0,:-1].values.reshape(1,-1))
-shap.force_plot(explainer.expected_value[0], shap_values[0], link='logit')  # repeat changing 0 for i in range(0, 5)
+# -- Dependence Contribution Plots
+# make plot.
+shap.dependence_plot('Scenicness.median_of_bayescore', shap_values[3], X_train, interaction_index = 'Wildness.median')
 
-# -- third trial
-shap.initjs()
-class_names = ['Low', 'Moderate', ' High', 'Outstanding']
-shap.summary_plot(shap_values, X_test, plot_type="bar", class_names= class_names, feature_names = ['Scenicness', 'Wildness'])
-
-# -- plot the global importance of each feature
-shap.plots.bar(explainer)
-
-# visualize the first prediction's explanation
-shap.force_plot(explainer.expected_value[0], shap_values[0])
-
-# fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
-# shap.summary_plot(shap_values)
-# # or 
-# shap.plots.beeswarm(shap_values)
-shap.summary_plot(shap_values[2], X_test)
-
-# fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
-shap.summary_plot(shap_values, X_test, plot_type="bar")
-
-shap_interaction_values = shap.TreeExplainer(model).shap_interaction_values(X_train)
-
-X_names = ['Scenicness.median','x_coord','y_coord']
-
-shap.summary_plot(shap_interaction_values, X_train, max_display=15, 
-                  feature_names = X_names,
-                  plot_type = 'dot')
-                  # plot_type="compact_dot")
-
-import shap  # package used to calculate Shap values
-
-# Create object that can calculate shap values
-explainer = shap.TreeExplainer(my_model)
-
-# calculate shap values. This is what we will plot.
-# Calculate shap_values for all of val_X rather than a single row, to have more data for plot.
-shap_values = explainer.shap_values(val_X)
-
-# Make plot. Index of [1] is explained in text below.
-shap.summary_plot(shap_values[1], val_X)                  
+# explainer = shap.TreeExplainer(model2)
+# shap_values = explainer.shap_values(X_sampled)
+# shap.summary_plot (shap_values, X_sampled, max_display=X_sampled.shape[1]) 
+# i=2
+# shap.force_plot(explainer.expected_value[0], shap_values[0][i], X_test.values[i], feature_names = X_test.columns)
+# 
+# 
+# fig.set_size_inches(18.5, 20.5)
+# fig.savefig("Z:\\R\\R projects\\LCA-GB\\models\\shap.png")
+# 
+# # -- first trial
+# shap.initjs()
+# shap_values = shap.TreeExplainer(model).shap_values(X_test)
+# shap.summary_plot(shap_values, X_test)
+# 
+# # -- second trial
+# scaler = StandardScaler()
+# scaled_data = scaler.fit_transform(X_test.iloc[:,:-1])
+# shap_values = explainer.shap_values(scaled_data.iloc[0,:-1].values.reshape(1,-1))
+# shap.force_plot(explainer.expected_value[0], shap_values[0], link='logit')  # repeat changing 0 for i in range(0, 5)
+# 
+# # -- third trial
+# shap.initjs()
+# class_names = ['Low', 'Moderate', ' High', 'Outstanding']
+# shap.summary_plot(shap_values, X_test, plot_type="bar", class_names= class_names, feature_names = ['Scenicness', 'Wildness'])
+# 
+# # -- plot the global importance of each feature
+# shap.plots.bar(explainer)
+# 
+# # visualize the first prediction's explanation
+# shap.force_plot(explainer.expected_value[0], shap_values[0])
+# 
+# # fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
+# # shap.summary_plot(shap_values)
+# # # or 
+# # shap.plots.beeswarm(shap_values)
+# shap.summary_plot(shap_values[2], X_test)
+# 
+# # fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
+# shap.summary_plot(shap_values, X_test, plot_type="bar")
+# 
+# shap_interaction_values = shap.TreeExplainer(model).shap_interaction_values(X_train)
+#     
+# X_names = ['Scenicness.median','x_coord','y_coord']
+# 
+# shap.summary_plot(shap_interaction_values, X_train, max_display=15, 
+#                   feature_names = X_names,
+#                   plot_type = 'dot')
+#                   # plot_type="compact_dot")
+#                   
+# import shap  # package used to calculate Shap values
+# 
+# # Create object that can calculate shap values
+# explainer = shap.TreeExplainer(my_model)
+# 
+# # calculate shap values. This is what we will plot.
+# # Calculate shap_values for all of val_X rather than a single row, to have more data for plot.
+# shap_values = explainer.shap_values(val_X)
+# 
+# # Make plot. Index of [1] is explained in text below.
+# shap.summary_plot(shap_values[1], val_X)                  
 ```

### 2022-08-19 paper3.Rmd
grid

@@
   mutate(ScenicQuality = factor(VS_46, levels = c("Low", "Moderate", "High", "Outstanding")),
-             Integrity = factor(VS_47, levels = c("Low", "Moderate", "High", "Outstanding")),
-             Character = factor(VS_48, levels = c("Low", "Moderate", "High", "Outstanding")),
-                Rarity = factor(VS_49, levels = c("Low", "Moderate", "High", "Outstanding")))
+         Consultant = ifelse(str_detect(VS_1a, "Bronwen Thomas"), "B", "A")) %>%
+             # Integrity = factor(VS_47, levels = c("Low", "Moderate", "High", "Outstanding")),
+             # Character = factor(VS_48, levels = c("Low", "Moderate", "High", "Outstanding")),
+             #    Rarity = factor(VS_49, levels = c("Low", "Moderate", "High", "Outstanding"))) %>%
+  dplyr::select(UID, CLS_1, CLS_2, CLS_3, CLS_4, Consultant, ScenicQuality)
 
@@
          Entropy = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
-         BayeScore = Votes %>% BayesRank(K = 10, α = .05))%>%
+         BayeScore = Votes %>% BayesRank(K = 10, α = .05),
+         Number = Votes %>% strsplit(",") %>% unlist() %>% length())%>%
   st_as_sf(coords = c("Lon","Lat"), crs = 4326) %>%
@@
   as_tibble() %>% 
-  st_as_sf() %>%
   rowwise() %>%
-  mutate(Scenicness.mean_of_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
-         Scenicness.median_of_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
-         Scenicness.entropy_of_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% entropy(),
-         Scenicness.number_of_all = Votes %>% strsplit(",") %>% unlist() %>% as.numeric() %>% na.omit() %>% length()) %>%
-  ungroup()
+  mutate(Scenicness.number_of_all = strsplit(Votes, ",") %>% unlist() %>% na.omit() %>% length(),
+         Scenicness.mean_of_all = strsplit(Votes, ",") %>% unlist() %>% as.numeric() %>% mean(),
+         Scenicness.median_of_all = strsplit(Votes, ",") %>% unlist() %>% as.numeric() %>% median(),
+         Scenicness.entropy_of_all = strsplit(Votes, ",") %>% unlist() %>% as.numeric() %>% entropy()) %>%
+  ungroup() %>%
+  st_as_sf()
 
@@
   as_tibble() %>%
-  st_as_sf() %>%
   rowwise() %>%
-  mutate(Scenicness.mean_of_median = Median %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
-         Scenicness.median_of_median = Median %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median()) %>%
-  ungroup()
+  mutate(Scenicness.mean_of_median = strsplit(Median, ",") %>% unlist() %>% as.numeric() %>% mean(),
+         Scenicness.median_of_median = strsplit(Median, ",") %>% unlist() %>% as.numeric() %>% median()) %>%
+  dplyr::select(-Median) %>%
+  ungroup() %>%
+  st_as_sf()
 
@@
   as_tibble() %>%
-  st_as_sf() %>%
   rowwise() %>%
-  mutate(Scenicness.mean_of_mean = Mean %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
-         Scenicness.median_of_mean = Mean %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median()) %>%
-  ungroup()
+  mutate(Scenicness.mean_of_mean = strsplit(Mean, ",") %>% unlist() %>% as.numeric() %>% mean(),
+         Scenicness.median_of_mean = strsplit(Mean, ",") %>% unlist() %>% as.numeric() %>% median()) %>%
+  dplyr::select(-Mean) %>%
+  ungroup() %>%
+  st_as_sf()
 
@@
   st_join(LANDMAP, .) %>%
-  aggregate(BayeScore~UID, data = ., paste0, collapse = ",") %>%
+  aggregate(cbind(BayeScore, Number, VR = BayeScore*Number)~UID, data = ., paste0, collapse = ",") %>%
   merge(LANDMAP, ., by = "UID", all = T) %>%
   as_tibble() %>%
-  st_as_sf() %>%
   rowwise() %>%
-  mutate(Scenicness.mean_of_bayescore = BayeScore %>% strsplit(",") %>% unlist() %>% as.numeric() %>% mean(),
-         Scenicness.median_of_bayescore = BayeScore %>% strsplit(",") %>% unlist() %>% as.numeric() %>% median(),
-         Consultant = ifelse(str_detect(VS_1a, "Bronwen Thomas"), "B", "A")) %>%
-  ungroup()
+  mutate(Scenicness.mean_of_bayescore = strsplit(BayeScore, ",") %>% unlist() %>% as.numeric() %>% mean(),
+         Scenicness.median_of_bayescore = strsplit(BayeScore, ",") %>% unlist() %>% as.numeric() %>% median(),
+         Score_multiplied_by_Number = strsplit(VR, ",") %>% unlist() %>% as.numeric() %>% sum(na.rm = TRUE),
+         Scenicness.weighted_bayescore =  Score_multiplied_by_Number/(Scenicness.number_of_all + 3) + 
+                                          3*Scenicness.mean_of_bayescore/(Scenicness.number_of_all + 3)) %>%
+  dplyr::select(-c(BayeScore, Number, VR, Score_multiplied_by_Number)) %>%
+  ungroup() %>%
+  st_as_sf()
 
-plot(LANDMAP["ScenicQuality"], border = 'NA', col = RColorBrewer::brewer.pal(4, "YlOrBr"), key.pos = 1)  
+# plot(LANDMAP["ScenicQuality"], border = 'NA', col = RColorBrewer::brewer.pal(4, "YlOrBr"), key.pos = 1)  
 ```
@@
                 quantiles = c(0.25, 0.75)) %>%
-  mutate(Wildness.iqr = q75 - q25) %>%
+  mutate(Wildness.mean = mean,
+         Wildness.median = median,
+         Wildness.iqr = q75 - q25,
+         Wildness.cv = coefficient_of_variation,
+         Wildness.variance = variance) %>%
+  dplyr::select(-c(mean, median, q25, q75, coefficient_of_variation, variance)) %>%
   bind_cols(LANDMAP, .) %>%
-  dplyr::select(UID, CLS_1, CLS_2, CLS_3, CLS_4, ScenicQuality, Consultant,
-                Votes, Scenicness.mean_of_all, Scenicness.median_of_all,
-                Scenicness.entropy_of_all, Scenicness.number_of_all,
-                Scenicness.mean_of_median, Scenicness.median_of_median,
-                Scenicness.mean_of_mean, Scenicness.median_of_mean,
-                Scenicness.mean_of_bayescore, Scenicness.median_of_bayescore,
-                Wildness.mean = mean, Wildness.median = median, Wildness.iqr, 
-                Wildness.cv = coefficient_of_variation, Wildness.variance = variance) %>%
   st_transform(crs = 27700)
@@
 landmap <- st_cast(LANDMAP, "POLYGON") %>%
-  dplyr::mutate(x = st_centroid(landmap$geometry) %>% st_coordinates() %>% .[, 1],
-                y = st_centroid(landmap$geometry) %>% st_coordinates() %>% .[, 2])
+  dplyr::mutate(UID = row_number(),
+                x_coord = st_centroid(landmap$geometry) %>% st_coordinates() %>% .[, 1],
+                y_coord = st_centroid(landmap$geometry) %>% st_coordinates() %>% .[, 2])
 summary(landmap)
+```
+
+Separate the consultant and then split train-test sets
+```{r}
+landmap_A = landmap %>% filter(Consultant == "A")
+landmap_B = landmap %>% filter(Consultant == "B")
+
+library(rsample)
+
+# Set the random number stream using `set.seed()` so that the results can be 
+# reproduced later. 
+set.seed(2046)
+
+# Save the split information for an 80/20 split of the data
+split <- landmap %>% 
+  filter(!is.na(ScenicQuality)) %>% 
+  initial_split(prop = 0.80, strata = ScenicQuality)
+
+train_set <- training(split) %>% st_drop_geometry() 
+ test_set <- testing(split) %>% st_drop_geometry() 
+dim(train_set)
+dim(test_set)
+
+X_train <- train_set %>% 
+  dplyr::select(Scenicness = Scenicness.median_of_bayescore,
+                Wildness = Wildness.median, x_coord, y_coord)
+ X_test <-  test_set %>% 
+   dplyr::select(Scenicness = Scenicness.median_of_bayescore,
+                 Wildness = Wildness.median, x_coord, y_coord)
+
+y_train <- train_set %>% dplyr::select(ScenicQuality)
+ y_test <-  test_set %>% dplyr::select(ScenicQuality)  
 ```
@@
 ```
-
 ```{python}
-import geopandas as gpd
-import pandas as pd
-from sklearn.preprocessing import OrdinalEncoder
-from sklearn.compose import make_column_transformer
-
-landmap = r.landmap
-landmap = landmap.dropna(subset=['ScenicQuality'])
-
-print(landmap)
-landmap.describe()
-print(landmap['ScenicQuality'].value_counts())
-
-landmap.isnull().sum()
-
-landmap_A = landmap.query("Consultant in ['A']")
-landmap_B = landmap.query("Consultant in ['B']")
-
-print(landmap.shape)
-print(landmap_A.shape)
-print(landmap_B.shape)
-
-# def split_consultant(landmap):
-#   X = landmap.loc[:, ['Scenicness.mean', 'Scenicness.entropy', 'Wildness.median', 'Wildness.iqr']]
-#   y = landmap.loc[:, ['ScenicQuality']]
-# 
-#   # Instantiate the label encoder
-#   transformer = make_column_transformer(
-#     (OrdinalEncoder(categories=[['Low', 'Moderate', 'High', 'Outstanding']]), ['ScenicQuality'])
-#   )
-# 
-#   y = transformer.fit_transform(y)
-#   
-#   y = pd.DataFrame(y, columns = ['ScenicQuality'])
-#   return X, y
-# 
-# def EDA_plot(y, X):
-#   import pandas as pd
-#   # concatenating y and X along columns
-#   landmap_concat = pd.DataFrame(pd.concat([y.reset_index(drop=True), X.reset_index(drop=True)], axis=1))
-#   print(landmap_concat.shape)
-# 
-#   import seaborn as sns
-#   sns.set()
-#   sns.pairplot(landmap_concat, hue='ScenicQuality', palette='vlag', height=3)
-
-# split train/test set
-# X = landmap_B.loc[:, ['Scenicness.mean', 'Wildness.mean']]
-X = landmap_B.loc[:, ['Scenicness.median_of_bayescore', 'Wildness.median', 'x', 'y']]
-y = landmap_B.loc[:, ['ScenicQuality']]
-
-# from sklearn.preprocessing import OrdinalEncoder
-# OE = OrdinalEncoder()
-# y[['Scale']] = OE.fit_transform(y[['ScenicQuality']])
+# -- load in the train/test sets
+X_train = r.X_train
+y_train = r.y_train
+ X_test = r.X_test
+ y_test = r.y_test
 
@@
                 'Outstanding': int(3)}
-
-y[['ScenicQuality']] = y[['ScenicQuality']].replace(scale_mapper).astype('int')
-
-# X, y = split_consultant(landmap)
-print(y.shape, X.shape)
-X.head()
-y.head()
-# EDA_plot(y, X)
-
-# -- Create train/test split
-from sklearn.model_selection import train_test_split
-X_train, X_test, y_train, y_test = train_test_split(X, y, 
-                                                    test_size = 0.2, 
-                                                    random_state = 2046, 
-                                                    stratify = y)
+y_train[['ScenicQuality']] = y_train[['ScenicQuality']].replace(scale_mapper).astype('int')
+ y_test[['ScenicQuality']] = y_test[['ScenicQuality']].replace(scale_mapper).astype('int')
 ```
-
 ## Hyperparameter Tuning
@@
 # Consultant B
-params = {
-          'booster': 'gbtree',
+params = {'booster': 'gbtree',
         'num_round': 1434, #1433.996,
@@
       'eval_metric': 'merror',
-             'seed': 1995
-  }
+             'seed': 1995}
 
@@
 
-# model = xgb.train(space_eval(param_space, best_params),
+model = xgb.train(space_eval(param_space, best_params),
+                  dtrain,
+                  num_boost_round = 500,
+                  evals = [(dtrain, 'train'), (dtest, 'test')],
+                  early_stopping_rounds = 250)
+
+# model = xgb.train(params,
 #                   dtrain,
@@
 #                   evals = [(dtrain, 'train'), (dtest, 'test')],
-#                   early_stopping_rounds = 250)
-
-model = xgb.train(params,
-                  dtrain,
-                  num_boost_round = 500,
-                  evals = [(dtrain, 'train'), (dtest, 'test')],
-                  verbose_eval = 100,
-                  early_stopping_rounds = 251)
+#                   verbose_eval = 100,
+#                   early_stopping_rounds = 251)
 
@@
 y_pred = model.predict(dtest)
+y_pred = y_pred.astype(int)
+y_true = y_test.values.flatten()
 
 # -- evaluate model using test data
+# -- overall accuracy
 from sklearn.metrics import accuracy_score
-
-accuracy = accuracy_score(y_test, y_pred)
+y_actual = y_test.values.flatten()
+accuracy = accuracy_score(y_true, y_pred)
 print("Accuracy: %.2f%%" % (accuracy * 100.0))
 
+# -- Cohen's kappa 
 from sklearn.metrics import cohen_kappa_score
-kappa = cohen_kappa_score(y_pred, y_test)
+kappa = cohen_kappa_score(y_true, y_pred)
 print("Kappa: %.2f" % (kappa))
@@
 # -- save to JSON
-model.save_model("Z:\\R\\R projects\\LCA-GB\\models\\model_B.json")
+model.save_model("Z:\\R\\R projects\\LCA-GB\\models\\model.json")
 ```
@@
 from sklearn.preprocessing import StandardScaler
-from sklearn.metrics import plot_confusion_matrix
 
@@
 model = xgb.XGBClassifier()
-model.load_model("Z:\\R\\R projects\\LCA-GB\\models\\model_B.json")
+model.load_model("Z:\\R\\R projects\\LCA-GB\\models\\model.json")
 ```
 
-## Confusion matrix
+## Confusion matrix and its visualisation
 ```{python}
+from sklearn.metrics import confusion_matrix, ConfusionMatrixDisplay
+import matplotlib.pyplot as plt
+
+cm = confusion_matrix(y_test.values, y_pred)
 # -- view the confusion matrix
-class_names = ['Low', 'Moderate', 'High', 'Outstanding']
-disp = ConfusionMatrixDisplay.from_predictions(model, X_test, y_test, 
-                             display_labels = class_names, 
-                             cmap = plt.cm.Blues, 
-                             xticks_rotation = 'vertical')
+disp = ConfusionMatrixDisplay(confusion_matrix = cm, 
+                              display_labels = ['Low', 'Moderate', 'High', 'Outstanding'])
+disp.plot(cmap = plt.cm.Blues)
+plt.show(); plt.close()
 ```
 
-```{python}
+```{python plot}
+X =  r.landmap.loc[:, ['Scenicness.median_of_bayescore', 'Wildness.median', 'x', 'y']]
 # -- play shap
 explainer = shap.TreeExplainer(model)
-shap_values = explainer.shap_values(X_train, approximate=True)
+# X = landmap.loc[:, ['UID', 'Scenicness.median_of_bayescore', 'Wildness.median', 'x', 'y']]
+# The approximate argument set to True to fix some of the shap_values with NaN value
+# shape of shap_values is : (4, 1321, 4) means 4 classes, 1321 rows(observations), each 4 columns(features).
+shap_values = explainer.shap_values(X, approximate = True)
 
-shap.plots.beeswarm(shap_values, show=False, color_bar=False)
-plt.colorbar()
-plt.show()
+# shap.plots.beeswarm(shap_values, show = True, color_bar = False)
+# plt.colorbar()
+# plt.show()
 
-# shap.summary_plot(shap_values[1], 
-#                   X_train.values, 
-#                   feature_names = X_train.columns,
-#                   plot_size = (10, 4))
+# In the case of matplotlib version 3.5 above, the adjustment of colorbar legend was referred to https://github.com/slundberg/shap/issues/2493
+# summary plot for "CLASS 0" (Low scenic quality) 
+shap.summary_plot(shap_values[0],
+                  X_train.values,
+                  feature_names = ['Scenicness', 'Wildness', 'x_coord', 'y_coord'], #X_train.columns,
+                  show = False,
+                  # plot_type = 'compact_dot', # making the dots disappear
+                  plot_size = (50, 6))
+plt.gcf().axes[-1].set_aspect('auto')
+plt.tight_layout()
+plt.gcf().axes[-1].set_box_aspect(40) # smaller "box_aspect" value to make colorbar thicker
+plt.show(); plt.close()
+# plt.savefig('fig_tes1.png', bbox_inches='tight',dpi=300); plt.close()
+
+# X_names = ['pct_18_34','pct_white','pct_bach',
+#           'pct_no_car','log_pop_den','job_entropy','network_den',
+#           'ave_trip_dist','pct_share','x_coord','y_coord']
 # 
-# plt.gcf().axes[-1].set_aspect('auto')
-# plt.tight_layout()
-# As mentioned, smaller "box_aspect" value to make colorbar thicker
-# plt.gcf().axes[-1].set_box_aspect(200)
+# fig, ax = plt.subplots(figsize=(8, 6),dpi=160)
+# 
+# shap.summary_plot(shap_interaction_values, X_coords, max_display=15, 
+#                   feature_names = X_names,
+#                   plot_type="compact_dot")
+```
+## Spatial patterns of SHAP values
+```{python}
+# landmap = r.landmap
+X_vars = ['Scenicness','Wildness', 'x', 'y']
 
+fig, ax = plt.subplots(2,2,figsize=(15,15))
+
+for j in range(len(X_vars)):
+  j = 1
+    ax = ax.ravel()
+    X.plot(ax=ax[j],column = shap_values.values[:,j],legend=True,
+                 vmin=-0.8,vmax=0.8,cmap=shap.plots.colors.red_white_blue)
+
+    ax[j].set_title("SHAP for\n" + X_vars[j],fontsize=10)
+
+#fig.delaxes(ax[-1])
+
+```
+
+```{r}
+shap_values = py$shap_values
+shap_Low = data.frame(shap_values[[1]]) %>% transmute(SHAP_Scenicness_Low = X1, 
+                                                      SHAP_Wildness_Low = X2,
+                                                      SHAP_Location_Low = X3 + X4) 
+shap_Moderate = data.frame(shap_values[[2]]) %>% transmute(SHAP_Scenicness_Moderate = X1,
+                                                           SHAP_Wildness_Moderate = X2,
+                                                           SHAP_Location_Moderate= X3 + X4)
+shap_High = data.frame(shap_values[[3]]) %>% transmute(SHAP_Scenicness_High = X1,
+                                                       SHAP_Wildness_High = X2,
+                                                       SHAP_Location_High = X3 + X4)
+shap_Outstanding = data.frame(shap_values[[4]]) %>% transmute(SHAP_Scenicness_Outstanding = X1,
+                                                              SHAP_Wildness_Outstanding = X2,
+                                                              SHAP_Location_Outstanding = X3 + X4)
+
+landmap_shap = cbind(landmap, shap_Low, shap_Moderate, shap_High, shap_Outstanding)
+rm(list = c('shap_Low', 'shap_Moderate', 'shap_High', 'shap_Outstanding'))
+
+```
+## Visualising the spatial distributions of SHAP values for each covariate with each class
+```{r}
+library(ggplot2)
+
+theme_map <- function(...) {
+  theme_minimal() +
+    theme(
+      text = element_text(color = "#22211d"), # family = "Garamond", "Ubuntu Regular"
+      axis.line = element_blank(),
+      axis.text.x = element_blank(),
+      axis.text.y = element_blank(),
+      axis.ticks = element_blank(),
+      axis.title.x = element_blank(),
+      axis.title.y = element_blank(),
+      # panel.grid.minor = element_line(color = "#ebebe5", size = 0.2),
+      panel.grid.major = element_blank(), #element_line(color = "#ebebe5", size = 0.2),
+      panel.grid.minor = element_blank(),
+      plot.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA), 
+      panel.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA), 
+      legend.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA),
+      panel.border = element_blank(),
+      ...
+    )
+}
+
+f <- list()
+
+for (i in 1:12) {
+  f[[i]] <- ggplot() +
+    geom_sf(data = landmap_shap, aes(fill = !!sym(names(landmap_shap)[i+25])), colour = NA) +
+    theme_map() +
+    labs(x = NULL, 
+         y = NULL, 
+         title = names(landmap_shap)[i+25]) + # stringr::str_c("SHAP for ", collapse = " ")
+    theme(legend.position = "bottom",
+          title = element_text(size = 15)) +
+    scale_fill_gradient2(low = ("#CB181D"),
+                         mid = "white",
+                         name = NULL,
+                         high = ("#2171B5"),
+                         midpoint = 0,
+                         space = "Lab",
+                         na.value = "white",
+                         guide = guide_colourbar(title.position = "top",
+                                                 # title.hjust = 0.5,
+                                                 title.vjust = 0.8,
+                                                 barwidth = 15))
+}
+
+## -- multiplot function
+multiplot <- function(plot.list, file, cols=3, layout=NULL) {
+	library(grid)
+	numPlots = length(plot.list)
+	if (is.null(layout)) {
+		layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+						ncol = cols, nrow = ceiling(numPlots/cols))
+	}
+	if (numPlots==1) {
+		print(plots[[1]])
+	} else {
+	  grid.newpage()
+	  pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+	  for (i in 1:numPlots) {
+	  	matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+	  	print(plot.list[[i]], vp = viewport(layout.pos.row = matchidx$row,
+	  									layout.pos.col = matchidx$col))
+	  }
+	}
+}
+
+## map together
+setwd("Z:\\R\\R projects\\LCA-GB\\figures")
+png(filename = "SHAP_values.png", w = 10*2, h = 6*2, units = "in", res = 300)
+par(mar = c(0,0,0,0))
+multiplot(list(f[1], f[2], f[3], f[4], 
+               f[5], f[6], f[7], f[8], 
+               f[9], f[10], f[11], f[12]), cols = 4)
+dev.off()
 ```

### 2022-08-26 paper3.Rmd
grid 1 km

@@
 X_train <- train_set %>% 
-  dplyr::select(Scenicness = Scenicness.median_of_bayescore,
+  dplyr::select(Scenicness = Scenicness.weighted_bayescore,
                 Wildness = Wildness.median, x_coord, y_coord)
  X_test <-  test_set %>% 
-   dplyr::select(Scenicness = Scenicness.median_of_bayescore,
+   dplyr::select(Scenicness = Scenicness.weighted_bayescore,
                  Wildness = Wildness.median, x_coord, y_coord)
@@
 repl_python() 
+```
 
-# reticulate::py_config() # why numpy_version couldn't be found?
-# use_python("Z:\\R\\R projects\\LCA-GB\\my_env_Python_3.9.13\\Scripts\\python.exe")
-# 
-# # use_virtualenv("Z:\\R\\R projects\\LCA-GB\\my_env_Python_3.9.13", required = TRUE)
-# 
-# install.packages("usethis")
-# usethis::edit_r_environ()
-# usethis::edit_r_profile()
-```
 ```{python}
@@
 
-model = xgb.train(space_eval(param_space, best_params),
+# model = xgb.train(space_eval(param_space, best_params),
+#                   dtrain,
+#                   num_boost_round = 500,
+#                   evals = [(dtrain, 'train'), (dtest, 'test')],
+#                   early_stopping_rounds = 250)
+
+model = xgb.train(params,
                   dtrain,
@@
                   evals = [(dtrain, 'train'), (dtest, 'test')],
-                  early_stopping_rounds = 250)
-
-# model = xgb.train(params,
-#                   dtrain,
-#                   num_boost_round = 500,
-#                   evals = [(dtrain, 'train'), (dtest, 'test')],
-#                   verbose_eval = 100,
-#                   early_stopping_rounds = 251)
+                  verbose_eval = 100,
+                  early_stopping_rounds = 251)
 
@@
 ```{python plot}
-X =  r.landmap.loc[:, ['Scenicness.median_of_bayescore', 'Wildness.median', 'x', 'y']]
+X =  r.landmap.loc[:, ['Scenicness.median_of_bayescore', 'Wildness.median', 'x_coord', 'y_coord']]
 # -- play shap
 explainer = shap.TreeExplainer(model)
-# X = landmap.loc[:, ['UID', 'Scenicness.median_of_bayescore', 'Wildness.median', 'x', 'y']]
+
 # The approximate argument set to True to fix some of the shap_values with NaN value
@@
 
-# In the case of matplotlib version 3.5 above, the adjustment of colorbar legend was referred to https://github.com/slundberg/shap/issues/2493
+# In the case of matplotlib version 3.5 above, the adjustment of colorbar legend was referred to https://stackoverflow.com/questions/70461753/shap-the-color-bar-is-not-displayed-in-the-summary-plot/
 # summary plot for "CLASS 0" (Low scenic quality)
-shap.summary_plot(shap_values[0],
-                  X_train.values,
-                  feature_names = ['Scenicness', 'Wildness', 'x_coord', 'y_coord'], #X_train.columns,
-                  show = False,
+
+# landmap = r.landmap
+X_vars = ['Scenicness','Wildness', 'x_coord', 'y_coord']
+
+import matplotlib.pyplot as plt
+
+fig, ax = plt.subplots(2,2,figsize=(15,15))
+
+for j in range(len(shap_values)):
+    ax = ax.ravel()
+    # X.plot(ax=ax[j],column = shap_values.values[:,j],legend=True,
+    #              vmin=-0.8,vmax=0.8,cmap=shap.plots.colors.red_white_blue)
+    fig[j] = shap.summary_plot(shap_values[1],
+                              X.values,
+                              feature_names = X_vars,
+                              show = False),
+                              # plot_type = 'compact_dot', # making the dots disappear
+                              plot_size = (50, 6))
+             plt.gcf().axes[-1].set_aspect('auto')
+             plt.tight_layout()
+             plt.gcf().axes[-1].set_box_aspect(40)
+    ax[j].set_title("SHAP for\n" + X_vars[j],fontsize=10)
+
+fig.delaxes(ax[-1])
+
+shap.summary_plot(shap_values[1],
+                  X.values,
+                  feature_names = ['Scenicness', 'Wildness', 'x_coord', 'y_coord'], show = False),
                   # plot_type = 'compact_dot', # making the dots disappear
@@
 # plt.savefig('fig_tes1.png', bbox_inches='tight',dpi=300); plt.close()
-
-# X_names = ['pct_18_34','pct_white','pct_bach',
-#           'pct_no_car','log_pop_den','job_entropy','network_den',
-#           'ave_trip_dist','pct_share','x_coord','y_coord']
-# 
-# fig, ax = plt.subplots(figsize=(8, 6),dpi=160)
-# 
-# shap.summary_plot(shap_interaction_values, X_coords, max_display=15, 
-#                   feature_names = X_names,
-#                   plot_type="compact_dot")
 ```
 ## Spatial patterns of SHAP values
-```{python}
-# landmap = r.landmap
-X_vars = ['Scenicness','Wildness', 'x', 'y']
-
-fig, ax = plt.subplots(2,2,figsize=(15,15))
-
-for j in range(len(X_vars)):
-  j = 1
-    ax = ax.ravel()
-    X.plot(ax=ax[j],column = shap_values.values[:,j],legend=True,
-                 vmin=-0.8,vmax=0.8,cmap=shap.plots.colors.red_white_blue)
-
-    ax[j].set_title("SHAP for\n" + X_vars[j],fontsize=10)
-
-#fig.delaxes(ax[-1])
-
-```
-
 ```{r}
@@
 dev.off()
+```
+# Change of support (to 1 km grid)
+```{r}
+# -- download the administrative region for Great Britain from GADM
+UK <- raster::getData('GADM', country = 'GBR', level = 0) |>
+  as('sf') |>
+  # filter(!NAME_1=='Northern Ireland') |>
+  st_transform(crs = 27700)
+plot(UK)
+
+# -- download OSGB 1 km grid from github and covert its coordinate system to 27700
+grid_1km <- "https://raw.githubusercontent.com/charlesroper/OSGB_Grids/master/GeoJSON/OSGB_Grid_1km.geojson" |>
+  geojsonio::geojson_read(what = "sp") |>
+  as("sf") |>
+  st_transform(crs = 27700)
+
+st_crs(grid_1km) == st_crs(sc)
+
+# -- spatial join OSGB 1 km grid with Scenic-Or-Not data
+test <- st_join(grid_1km, sc)
+
+test1 <- test |>
+  aggregate(cbind(ID,
+                  `Geograph URI`,
+                  Votes,
+                  Number,
+                  Mean,
+                  Median,
+                  BayeScore,
+                  Entropy,
+                  Variance)~PLAN_NO, data = ., paste0, collapse = ",") %>%
+  merge(LANDMAP, ., by = "UID", all = T) %>%
+  as_tibble()
+
+plot(UK['NAME_0']) +
+  plot(test[is.na(test$ID),'Mean'])
+
+grid_1km <- st_join(grid_1km, sc, join = st_within, left = TRUE) |>
+  dplyr::select(PLAN_NO, 
+                `Scenic-Or-Not ID` = ID,
+                `Geograph URI`,
+                Votes,
+                Number,
+                Mean,
+                Median,
+                BayeScore,
+                Entropy,
+                Variance)
+
+grid_1km <- st_transform(grid_1km, crs = st_crs(Wildness)) %>%
+  exact_extract(Wildness, ., 
+                fun = c('mean', 'median', 'coefficient_of_variation', 'variance', 'quantile'), 
+                quantiles = c(0.25, 0.75)) |>
+  mutate(Wildness.mean = mean,
+         Wildness.median = median,
+         Wildness.iqr = q75 - q25,
+         Wildness.cv = coefficient_of_variation,
+         Wildness.variance = variance) |>
+  dplyr::select(-c(mean, median, q25, q75, coefficient_of_variation, variance)) |>
+  bind_cols(grid_1km, .) |>
+  st_transform(crs = 27700)
+
 ```

### 2022-09-01 paper3.Rmd
2nd Sep

@@
 test <- st_join(grid_1km, sc) 
+library(dplyr)
 
 test1 <- test |>
-  aggregate(cbind(ID,
-                  `Geograph URI`,
-                  Votes,
-                  Number,
-                  Mean,
-                  Median,
-                  BayeScore,
-                  Entropy,
-                  Variance)~PLAN_NO, data = ., paste0, collapse = ",") %>%
-  merge(LANDMAP, ., by = "UID", all = T) %>%
+  st_drop_geometry() |>
+  group_by(PLAN_NO) |>
+  summarise(ID = paste2(ID, collapse = ', '),
+            `Geograph URI` = paste(`Geograph URI`, collapse = ', '),
+            Votes = paste(Votes, collapse = ', '),
+            Number = paste(Number, collapse = ', '),
+            Mean = paste(Mean, collapse = ', '),
+            Median = paste(Median, collapse = ', '),
+            BayeScore = paste(BayeScore, collapse = ', '),
+            Entopy = paste(Entropy, collapse = ', ')) |>
+  ungroup() |>
+  rowwise() |>
+
+test2 <- test1 |>  
+  mutate(Photos = strsplit(ID, ',') |> unlist() |> !is.na() |> length())
+
+strsplit(",") %>% unlist() %>% as.numeric() %>% mean()
+
+data %>% 
+      group_by(foo) %>% 
+      mutate(bars_by_foo = paste0(bar, collapse = ""))
+
+# library(dplyr)
+# df %>% 
+#   group_by_(~page, ~passage, ~person) %>%
+#   summarize_(text=~paste(text, collapse=' '))
+
+  aggregate(#cbind(# ID,
+                  # `Geograph URI`,
+                  # Votes,
+                  # Number,
+                  # Mean,
+                  # Median,
+                  # BayeScore,
+                  # Entropy,
+                  Variance~PLAN_NO, data = ., FUN = paste, collapse = ",") %>%
+  merge(grid_1km, ., by = "PLAN_NO", all = T) |>
   as_tibble()

### 2022-09-19 paper3.Rmd
9/16

--- 2022-08-26 paper3.Rmd
@@
 # Save the split information for an 80/20 split of the data
-split <- landmap %>% 
+split <- landmap_B %>% 
   filter(!is.na(ScenicQuality)) %>% 
@@
   dplyr::select(Scenicness = Scenicness.weighted_bayescore,
-                Wildness = Wildness.median, x_coord, y_coord)
+                Wildness = Wildness.median, 
+                x_coord, y_coord)
  X_test <-  test_set %>% 
    dplyr::select(Scenicness = Scenicness.weighted_bayescore,
-                 Wildness = Wildness.median, x_coord, y_coord)
+                 Wildness = Wildness.median, 
+                 x_coord, y_coord)
 
@@
 ```{python plot}
-X =  r.landmap.loc[:, ['Scenicness.median_of_bayescore', 'Wildness.median', 'x_coord', 'y_coord']]
+X =  r.landmap.loc[:, ['Scenicness.weighted_bayescore', 'Wildness.median', 'x_coord', 'y_coord']]
 # -- play shap
@@
 ```{r}
-# -- download the administrative region for Great Britain from GADM
-UK <- raster::getData('GADM', country = 'GBR', level = 0) |>
-  as('sf') |>
-  # filter(!NAME_1=='Northern Ireland') |>
-  st_transform(crs = 27700)
-plot(UK)
-
 # -- download OSGB 1 km grid from github and covert its coordinate system to 27700
@@
 
-# -- spatial join OSGB 1 km grid with Scenic-Or-Not data
-test <- st_join(grid_1km, sc)
+# -- spatial join OSGB 1 km grid with Scenic-Or-Not data for the sake of understanding the distribution of Scenic-Or-Not ratings
 
-test1 <- test |>
-  aggregate(cbind(ID,
-                  `Geograph URI`,
-                  Votes,
-                  Number,
-                  Mean,
-                  Median,
-                  BayeScore,
-                  Entropy,
-                  Variance)~PLAN_NO, data = ., paste0, collapse = ",") %>%
-  merge(LANDMAP, ., by = "UID", all = T) %>%
-  as_tibble()
+# grid_1km <- st_join(grid_1km, sc) |>
+#   st_drop_geometry() |>
+#   group_by(PLAN_NO) |>
+#   summarise(ID = paste(ID, collapse = ','),
+#             `Geograph URI` = paste(`Geograph URI`, collapse = ','),
+#             Votes = paste(Votes, collapse = ','),
+#             Number = paste(Number, collapse = ','),
+#             Mean = paste(Mean, collapse = ','),
+#             Median = paste(Median, collapse = ','),
+#             BayeScore = paste(BayeScore, collapse = ','),
+#             Entopy = paste(Entropy, collapse = ',')) |>
+#   rowwise() |>
+#   mutate(Photos = ifelse(ID == 'NA', 0, strsplit(ID, ',') %>% unlist() %>% length())) |>
+#   ungroup() %>%
+#   left_join(grid_1km, ., by = c('PLAN_NO' = 'PLAN_NO'))
 
-plot(UK['NAME_0']) +
-  plot(test[is.na(test$ID),'Mean'])
+grid_1km <- st_join(grid_1km, sc) %>%
+  aggregate(cbind(BayeScore, Number, VR = BayeScore*Number)~PLAN_NO, data = ., paste0, collapse = ',') %>%
+  merge(grid_1km, ., by = 'PLAN_NO', all = T) |>
+  st_drop_geometry() |>
+  rowwise() |>    
+  mutate(Scenicness.number_of_all = strsplit(Number, ',') %>% unlist() %>% as.numeric() %>% sum(na.rm = TRUE),
+         Scenicness.mean_of_bayescore = strsplit(BayeScore, ',') %>% unlist() %>% as.numeric() %>% mean(),
+         Scenicness.median_of_bayescore = strsplit(BayeScore, ',') %>% unlist() %>% as.numeric() %>% median(),
+         Score_multiplied_by_Number = strsplit(VR, ',') %>% unlist() %>% as.numeric() %>% sum(na.rm = TRUE),
+         Scenicness.weighted_bayescore =  Score_multiplied_by_Number/(Scenicness.number_of_all + 3) + 
+                                          3*Scenicness.mean_of_bayescore/(Scenicness.number_of_all + 3)) |>
+  dplyr::select(-c(BayeScore, Number, VR, Score_multiplied_by_Number, Scenicness.number_of_all)) |>
+  ungroup() %>%
+  left_join(grid_1km, ., by = c('PLAN_NO' = 'PLAN_NO'))
 
-grid_1km <- st_join(grid_1km, sc, join = st_within, left = TRUE) |>
-  dplyr::select(PLAN_NO, 
-                `Scenic-Or-Not ID` = ID,
-                `Geograph URI`,
-                Votes,
-                Number,
-                Mean,
-                Median,
-                BayeScore,
-                Entropy,
-                Variance)
+# -- extract raster values of wildness quality based on 1 km OSGB grid 
+library(exactextractr)
+
+Wildness <- raster('Z:\\R\\R projects\\LCA-GB\\James\\mce_eq256')
+
+st_crs(Wildness) == st_crs(grid_1km)
 
@@
          Wildness.variance = variance) |>
-  dplyr::select(-c(mean, median, q25, q75, coefficient_of_variation, variance)) |>
+  dplyr::select(-c(mean, median, q25, q75, coefficient_of_variation, variance)) %>%
   bind_cols(grid_1km, .) |>
@@
 
+# -- calculate the centroid coordinate for each grid cell
+grid_1km <- mutate(grid_1km, 
+                   x_coord = st_centroid(grid_1km$geometry) %>% st_coordinates() %>% .[, 1],
+                   y_coord = st_centroid(grid_1km$geometry) %>% st_coordinates() %>% .[, 2])
+
+# -- download the administrative region for Great Britain from GADM for subseting the 1 km grid
+GB <- raster::getData('GADM', country = 'GBR', level = 1) |>
+  as('sf') |>
+  filter(!NAME_1=='Northern Ireland') |>
+  st_union() |>
+  st_transform(crs = 27700)
+plot(GB)
+
+grid_1km <- grid_1km[GB, ]
+```
+# prepare environment for running python
+```{r}
+Sys.setenv("RETICULATE_PYTHON" = ".\\.venv\\Scripts\\python.exe")
+# Sys.getenv()
+Sys.which("Python")
+
+# py_discover_config()
+
+library(reticulate)
+# -- creates an interactive Python console, REPL (Read, Eval, Print Loop), within R.
+repl_python() 
+```
+
+# SHAP value
+```{python plot}
+import shap
+
+X =  r.grid_1km.loc[:, ['Scenicness.weighted_bayescore', 'Wildness.median', 'x_coord', 'y_coord']]
+# -- play shap
+explainer = shap.TreeExplainer(model)
+
+# The approximate argument set to True to fix some of the shap_values with NaN value
+# shape of shap_values is : (4, 1321, 4) means 4 classes, 1321 rows(observations), each 4 columns(features).
+shap_values = explainer.shap_values(X, approximate = True)
+
+# shap.plots.beeswarm(shap_values, show = True, color_bar = False)
+# plt.colorbar()
+# plt.show()
+
+# In the case of matplotlib version 3.5 above, the adjustment of colorbar legend was referred to https://stackoverflow.com/questions/70461753/shap-the-color-bar-is-not-displayed-in-the-summary-plot/
+# summary plot for "CLASS 0" (Low scenic quality)
+
+# landmap = r.landmap
+
+# X_vars = ['Scenicness','Wildness', 'x_coord', 'y_coord']
+# 
+# import matplotlib.pyplot as plt
+# 
+# fig, ax = plt.subplots(2,2,figsize=(15,15))
+# 
+# for j in range(len(shap_values)):
+#     ax = ax.ravel()
+#     # X.plot(ax=ax[j],column = shap_values.values[:,j],legend=True,
+#     #              vmin=-0.8,vmax=0.8,cmap=shap.plots.colors.red_white_blue)
+#     fig[j] = shap.summary_plot(shap_values[1],
+#                               X.values,
+#                               feature_names = X_vars,
+#                               show = False),
+#                               # plot_type = 'compact_dot', # making the dots disappear
+#                               plot_size = (50, 6))
+#              plt.gcf().axes[-1].set_aspect('auto')
+#              plt.tight_layout()
+#              plt.gcf().axes[-1].set_box_aspect(40)
+#     ax[j].set_title("SHAP for\n" + X_vars[j],fontsize=10)
+#     
+# fig.delaxes(ax[-1])
+# 
+# 
+# shap.summary_plot(shap_values[1],
+#                   X.values,
+#                   feature_names = ['Scenicness', 'Wildness', 'x_coord', 'y_coord'], show = False),
+#                   # plot_type = 'compact_dot', # making the dots disappear
+#                   plot_size = (50, 6))
+# plt.gcf().axes[-1].set_aspect('auto')
+# plt.tight_layout()
+# plt.gcf().axes[-1].set_box_aspect(40) # smaller "box_aspect" value to make colorbar thicker
+# plt.show(); plt.close()
+# plt.savefig('fig_tes1.png', bbox_inches='tight',dpi=300); plt.close()
+```
+## Spatial patterns of SHAP values
+```{r}
+shap_values = py$shap_values
+shap_Low = data.frame(shap_values[[1]]) %>% transmute(SHAP_Scenicness_Low = X1, 
+                                                      SHAP_Wildness_Low = X2,
+                                                      SHAP_Location_Low = X3 + X4) 
+shap_Moderate = data.frame(shap_values[[2]]) %>% transmute(SHAP_Scenicness_Moderate = X1,
+                                                           SHAP_Wildness_Moderate = X2,
+                                                           SHAP_Location_Moderate= X3 + X4)
+shap_High = data.frame(shap_values[[3]]) %>% transmute(SHAP_Scenicness_High = X1,
+                                                       SHAP_Wildness_High = X2,
+                                                       SHAP_Location_High = X3 + X4)
+shap_Outstanding = data.frame(shap_values[[4]]) %>% transmute(SHAP_Scenicness_Outstanding = X1,
+                                                              SHAP_Wildness_Outstanding = X2,
+                                                              SHAP_Location_Outstanding = X3 + X4)
+
+grid_1km_shap = cbind(grid_1km, shap_Low, shap_Moderate, shap_High, shap_Outstanding)
+rm(list = c('shap_Low', 'shap_Moderate', 'shap_High', 'shap_Outstanding'))
+```
+## plotting the spatial distribution of SHAP values
+```{r}
+library(ggplot2)
+
+f2 <- list()
+
+for (i in 1:12) {
+  f2[[i]] <- ggplot() +
+    geom_sf(data = grid_1km_shap, aes(fill = !!sym(names(grid_1km_shap)[i+11])), colour = NA) +
+    theme_map() +
+    labs(x = NULL, 
+         y = NULL, 
+         title = names(grid_1km_shap)[i+11]) + # stringr::str_c("SHAP for ", collapse = " ")
+    theme(legend.position = "bottom",
+          title = element_text(size = 15)) +
+    scale_fill_gradient2(low = ("#CB181D"),
+                         mid = "white",
+                         name = NULL,
+                         high = ("#2171B5"),
+                         midpoint = 0,
+                         space = "Lab",
+                         na.value = "white",
+                         guide = guide_colourbar(title.position = "top",
+                                                 # title.hjust = 0.5,
+                                                 title.vjust = 0.8,
+                                                 barwidth = 15))
+}
+
+setwd("Z:\\R\\R projects\\LCA-GB\\figures")
+png(filename = "SHAP_values_4.png", w = 10*2, h = 6*2, units = "in", res = 300)
+par(mar = c(0,0,0,0))
+multiplot(list(f2[1], f2[2], f2[3], f2[4], 
+               f2[5], f2[6], f2[7], f2[8], 
+               f2[9], f2[10], f2[11], f2[12]), cols = 4)
+dev.off()
 ```

### 2022-12-02 LCA-GB.Rmd
20221202

@@
 
-```{r}
-# Model evaluation
-#predicted = predict(xgb_tune, x_test)
-# Predict outcomes with the test data
-Test %>% dplyr::select(Sce, Abs, Nat, Rem, Rug) %>%
-  st_drop_geometry() %>% 
-  as.matrix() %>% 
-  xgb.DMatrix() %>%
-  predict(xgb_tune, ., reshape = T) %>%
-  plyr::mapvalues(from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")) %>%
-  data.frame(predict = .) %>%
-  data.frame(actual = plyr::mapvalues(Test$SQ, from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")),
-             predict = .) ->
-  result
-
-# confusion matrix
-confusionMatrix(result$actual, result$predict) %>% print()
-
-# Reference
-# https://datascienceplus.com/extreme-gradient-boosting-with-r/
-
-hex.2k$SQ.pred <-
-  hex.2k %>%
-  st_drop_geometry() %>%
-  dplyr::select(Sce, Abs, Nat, Rem, Rug) %>%
-  predict(xgb_tune, ., reshape = T) %>%
-  plyr::mapvalues(from = c(1, 2, 3, 4), to = c("Low", "Moderate", "High", "Outstanding")) %>%
-  as.factor()
-```
-
-# Visualisation
-
-```{r}
-theme_map <- function(...) {
-  theme_minimal() +
-    theme(
-      text = element_text(family = "Arial", color = "#22211d"), #"Ubuntu Regular"
-      axis.line = element_blank(),
-      axis.text.x = element_blank(),
-      axis.text.y = element_blank(),
-      axis.ticks = element_blank(),
-      axis.title.x = element_blank(),
-      axis.title.y = element_blank(),
-      # panel.grid.minor = element_line(color = "#ebebe5", size = 0.2),
-      panel.grid.major = element_blank(), #element_line(color = "#ebebe5", size = 0.2),
-      panel.grid.minor = element_blank(),
-      plot.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA), 
-      panel.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA), 
-      legend.background = element_blank(), #element_rect(fill = "#f5f5f2", color = NA),
-      panel.border = element_blank(),
-      ...
-    )
-}
-
-p <- ggplot(hex.2k) +
-  # municipality polygons
-  geom_sf(aes(fill = SQ.pred), color = NA) +
-  theme_map() +
-  labs(x = NULL, y = NULL, 
-       title = "Predicted scenic quality") +
-  theme(legend.position = "bottom",
-        title =element_text(size = 15)) +
-  scale_fill_manual(
-  # in manual scales, one has to define colors, well, manually
-  # I can directly access them using viridis' magma-function
-  values = RColorBrewer::brewer.pal(4, "YlOrBr"), #rev(magma(9)),
-  #breaks = brks_scale,
-  name = "Quality levels",
-  drop = FALSE,
-  labels = c("Low", "Moderate", "High", "Outstanding"),
-  guide = guide_legend(
-    direction = "horizontal",
-    keyheight = unit(6, units = "mm"),
-    keywidth = unit(30 / length(labels), units = "mm"),
-    title.position = 'top',
-    title.theme = element_text(size = 15),
-    # I shift the labels around, the should be placed 
-    # exactly at the right end of each legend key
-    # title.hjust = -0.9,
-    label.hjust = 0.2,
-    label.theme = element_text(size = 12),
-    nrow = 1,
-    byrow = T,
-    # also the guide needs to be reversed
-    reverse = F,
-    label.position = "bottom")
-  )
-
-plot(test[, "SQ"])
-```
+\`\`\`{r} \# Model evaluation #predicted = predict(xgb_tune, x_test) \# Predict outcomes with the test data Test %\>% dplyr::select(Sce, Abs,

### 2022-12-02 paper3.Rmd
20221202

@@
                 fun = c('mean', 'median', 'coefficient_of_variation', 'variance', 'quantile'), 
-                quantiles = c(0.25, 0.75)) %>%
+                quantiles = c(0.25, 0.75)) |>
   mutate(Wildness.mean = mean,
@@
          Wildness.cv = coefficient_of_variation,
-         Wildness.variance = variance) %>%
+         Wildness.variance = variance) |>
   dplyr::select(-c(mean, median, q25, q75, coefficient_of_variation, variance)) %>%
-  bind_cols(LANDMAP, .) %>%
+  bind_cols(LANDMAP, .) |>
   st_transform(crs = 27700)
 
-landmap <- st_cast(LANDMAP, "POLYGON") %>%
+landmap <- st_cast(LANDMAP, "POLYGON") |>
   dplyr::mutate(UID = row_number(),
@@
 summary(landmap)
+
+plot(st_centroid(landmap)[, 1])
+plot(d[, 'centroids'], add = T, col = 'red', pch = 19)
+
+test <- st_centroid(landmap)
+  st_centroid() %>% 
+  # this is the crs from d, which has no EPSG code:
+  st_transform(., '+proj=longlat +ellps=GRS80 +no_defs') %>%
+  # since you want the centroids in a second geometry col:
+  st_geometry()
 ```
@@
 Separate the consultant and then split train-test sets
+
 ```{r}
@@
   dplyr::select(Scenicness = Scenicness.weighted_bayescore,
-                Wildness = Wildness.median, 
+                Wildness = Wildness.mean, 
                 x_coord, y_coord)
@@
    dplyr::select(Scenicness = Scenicness.weighted_bayescore,
-                 Wildness = Wildness.median, 
+                 Wildness = Wildness.mean,
                  x_coord, y_coord)
@@
                 'Outstanding': int(3)}
+
 y_train[['ScenicQuality']] = y_train[['ScenicQuality']].replace(scale_mapper).astype('int')
@@
 #        'eval_metric': 'merror',
-#               'seed': 2046
-#     }
+#               'seed': 2046}
 
@@
 ## Confusion matrix and its visualisation
+
 ```{python}
@@
                                plot_size = (50, 6))
+
              plt.gcf().axes[-1].set_aspect('auto')
              plt.tight_layout()
-             plt.gcf().axes[-1].set_box_aspect(40)
+             plt.gcf().axes[-1].set_box_aspect(40) # smaller "box_aspect" value to make colorbar thicker
     ax[j].set_title("SHAP for\n" + X_vars[j],fontsize=10)
@@
 
-shap.summary_plot(shap_values[1],
-                  X.values,
-                  feature_names = ['Scenicness', 'Wildness', 'x_coord', 'y_coord'], show = False),
-                  # plot_type = 'compact_dot', # making the dots disappear
-                  plot_size = (50, 6))
-plt.gcf().axes[-1].set_aspect('auto')
-plt.tight_layout()
-plt.gcf().axes[-1].set_box_aspect(40) # smaller "box_aspect" value to make colorbar thicker
-plt.show(); plt.close()
+# 
+# shap.summary_plot(shap_values[1],
+#                   X.values,
+#                   feature_names = ['Scenicness', 'Wildness', 'x_coord', 'y_coord'], show = False),
+#                   # plot_type = 'compact_dot', # making the dots disappear
+#                   plot_size = (50, 6))
+# plt.gcf().axes[-1].set_aspect('auto')
+# plt.tight_layout()
+# plt.gcf().axes[-1].set_box_aspect(40) 
+# plt.show(); plt.close()
 # plt.savefig('fig_tes1.png', bbox_inches='tight',dpi=300); plt.close()
 ```
+
 ## Spatial patterns of SHAP values
+
 ```{r}
@@
 ```
+
 ## Visualising the spatial distributions of SHAP values for each covariate with each class
+
 ```{r}
@@
 ```
+
 # Change of support (to 1 km grid)
+
 ```{r}
@@
 
+rm('Wildness')
+
 # -- calculate the centroid coordinate for each grid cell
@@
 
-# -- download the administrative region for Great Britain from GADM for subseting the 1 km grid
+# -- download the administrative region for Great Britain from GADM for sub-setting the 1 km grid
 GB <- raster::getData('GADM', country = 'GBR', level = 1) |>
@@
 grid_1km <- grid_1km[GB, ]
+
+# -- remove the grid cells that lack of scenicness and wildness measures
+grid_1km <- grid_1km[!is.na(grid_1km$Scenicness.weighted_bayescore), ] %>%
+  .[!is.na(grid_1km$Wildness.median), ]
 ```
+
 # prepare environment for running python
+
 ```{r}
@@
 # SHAP value
+
 ```{python plot}
@@
 shap_values = explainer.shap_values(X, approximate = True)
+```
 
-# shap.plots.beeswarm(shap_values, show = True, color_bar = False)
-# plt.colorbar()
-# plt.show()
+## Spatial patterns of SHAP values
 
-# In the case of matplotlib version 3.5 above, the adjustment of colorbar legend was referred to https://stackoverflow.com/questions/70461753/shap-the-color-bar-is-not-displayed-in-the-summary-plot/
-# summary plot for "CLASS 0" (Low scenic quality)
-
-# landmap = r.landmap
-
-# X_vars = ['Scenicness','Wildness', 'x_coord', 'y_coord']
-# 
-# import matplotlib.pyplot as plt
-# 
-# fig, ax = plt.subplots(2,2,figsize=(15,15))
-# 
-# for j in range(len(shap_values)):
-#     ax = ax.ravel()
-#     # X.plot(ax=ax[j],column = shap_values.values[:,j],legend=True,
-#     #              vmin=-0.8,vmax=0.8,cmap=shap.plots.colors.red_white_blue)
-#     fig[j] = shap.summary_plot(shap_values[1],
-#                               X.values,
-#                               feature_names = X_vars,
-#                               show = False),
-#                               # plot_type = 'compact_dot', # making the dots disappear
-#                               plot_size = (50, 6))
-#              plt.gcf().axes[-1].set_aspect('auto')
-#              plt.tight_layout()
-#              plt.gcf().axes[-1].set_box_aspect(40)
-#     ax[j].set_title("SHAP for\n" + X_vars[j],fontsize=10)
-#     
-# fig.delaxes(ax[-1])
-# 
-# 
-# shap.summary_plot(shap_values[1],
-#                   X.values,
-#                   feature_names = ['Scenicness', 'Wildness', 'x_coord', 'y_coord'], show = False),
-#                   # plot_type = 'compact_dot', # making the dots disappear
-#                   plot_size = (50, 6))
-# plt.gcf().axes[-1].set_aspect('auto')
-# plt.tight_layout()
-# plt.gcf().axes[-1].set_box_aspect(40) # smaller "box_aspect" value to make colorbar thicker
-# plt.show(); plt.close()
-# plt.savefig('fig_tes1.png', bbox_inches='tight',dpi=300); plt.close()
-```
-## Spatial patterns of SHAP values
 ```{r}
@@
 ```
+
 ## plotting the spatial distribution of SHAP values
+
 ```{r}
@@
 
-f2 <- list()
+f <- list()
 
 for (i in 1:12) {
-  f2[[i]] <- ggplot() +
+  f[[i]] <- ggplot() +
     geom_sf(data = grid_1km_shap, aes(fill = !!sym(names(grid_1km_shap)[i+11])), colour = NA) +
@@
 
+f[13] <- ggplot() +
+  geom_sf(data = grid_1km_shap, aes(fill = !!sym(names(grid_1km_shap)[4])), colour = NA) +
+  theme_map() +
+  labs(x = NULL, 
+       y = NULL, 
+       title = names(grid_1km_shap)[4]) + # stringr::str_c("SHAP for ", collapse = " ")
+  theme(legend.position = "bottom",
+        title = element_text(size = 15)) +
+  scale_fill_gradient2(low = ("#CB181D"),
+                       mid = "white",
+                       name = NULL,
+                       high = ("#2171B5"),
+                       midpoint = 0,
+                         space = "Lab",
+                         na.value = "white",
+                         guide = guide_colourbar(title.position = "top",
+                                                 # title.hjust = 0.5,
+                                                 title.vjust = 0.8,
+                                                 barwidth = 15))
+
+f[14] <- ggplot() +
+  geom_sf(data = grid_1km_shap, aes(fill = !!sym(names(grid_1km_shap)[6])), colour = NA) +
+  theme_map() +
+  labs(x = NULL, 
+       y = NULL, 
+       title = names(grid_1km_shap)[6]) + # stringr::str_c("SHAP for ", collapse = " ")
+  theme(legend.position = "bottom",
+        title = element_text(size = 15)) +
+  scale_fill_gradient2(low = ("#CB181D"),
+                       mid = "white",
+                       name = NULL,
+                       high = ("#2171B5"),
+                       midpoint = 0,
+                         space = "Lab",
+                         na.value = "white",
+                         guide = guide_colourbar(title.position = "top",
+                                                 # title.hjust = 0.5,
+                                                 title.vjust = 0.8,
+                                                 barwidth = 15))
+
+f[15] <- ggplot() +
+  geom_sf(data = grid_1km_shap, aes(fill = !!sym(names(grid_1km_shap)[10])), colour = NA) +
+  theme_map() +
+  labs(x = NULL, 
+       y = NULL, 
+       title = names(grid_1km_shap)[10]) + # stringr::str_c("SHAP for ", collapse = " ")
+  theme(legend.position = "bottom",
+        title = element_text(size = 15)) +
+  scale_fill_gradient2(low = ("#CB181D"),
+                       mid = "white",
+                       name = NULL,
+                       high = ("#2171B5"),
+                       midpoint = 0,
+                         space = "Lab",
+                         na.value = "white",
+                         guide = guide_colourbar(title.position = "top",
+                                                 # title.hjust = 0.5,
+                                                 title.vjust = 0.8,
+                                                 barwidth = 15))
+
+f[16] <- ggplot() +
+  geom_sf(data = grid_1km_shap, aes(fill = !!sym(names(grid_1km_shap)[11])), colour = NA) +
+  theme_map() +
+  labs(x = NULL, 
+       y = NULL, 
+       title = names(grid_1km_shap)[11]) + # stringr::str_c("SHAP for ", collapse = " ")
+  theme(legend.position = "bottom",
+        title = element_text(size = 15)) +
+  scale_fill_gradient2(low = ("#CB181D"),
+                       mid = "white",
+                       name = NULL,
+                       high = ("#2171B5"),
+                       midpoint = 0,
+                         space = "Lab",
+                         na.value = "white",
+                         guide = guide_colourbar(title.position = "top",
+                                                 # title.hjust = 0.5,
+                                                 title.vjust = 0.8,
+                                                 barwidth = 15))
+
 setwd("Z:\\R\\R projects\\LCA-GB\\figures")
-png(filename = "SHAP_values_4.png", w = 10*2, h = 6*2, units = "in", res = 300)
+png(filename = "SHAP_values_5.png", w = 10*2, h = 6*2, units = "in", res = 300)
 par(mar = c(0,0,0,0))
-multiplot(list(f2[1], f2[2], f2[3], f2[4], 
-               f2[5], f2[6], f2[7], f2[8], 
-               f2[9], f2[10], f2[11], f2[12]), cols = 4)
+multiplot(list(f[1], f[2], f[3], f[4], f[13],
+               f[5], f[6], f[7], f[8], f[14],
+               f[9], f[10], f[11], f[12], f[15]), cols = 5)
 dev.off()
+rm(c('i', 'f'))
 ```
+
+#### Waiting for deletion -- The XGBoost built-in feature importance
 
@@
 plt.title('Feature Importances')
-plt.barh(range(len(indices)), importances[indices], color='g', align='center')
+plt.barh(range(len(indices)), importances[indices], color = 'g', align = 'center')
 plt.yticks(range(len(indices)), [features[i] for i in indices])
@@
 
-```{python}
-# get just the explanations for the positive class
-# shap_values = shap_values[...,1]
-
-# shap.plots.bar(shap_values[...,1])
-```
+#### Waiting for deletion -- The dependence plots may not be used.
 
@@
 # shap_values[1] what does the index mean?
-shap.summary_plot(shap_values[1], X_train.values, feature_names=X_train.columns)
+shap.summary_plot(shap_values[1], X_train.values, feature_names = X_train.columns)
 # shap.summary_plot(shap_values, X_test, max_display = X_test.shape[1])
@@
 shap.dependence_plot('Scenicness.median_of_bayescore', shap_values[3], X_train, interaction_index = 'Wildness.median')
-
-# explainer = shap.TreeExplainer(model2)
-# shap_values = explainer.shap_values(X_sampled)
-# shap.summary_plot (shap_values, X_sampled, max_display=X_sampled.shape[1]) 
-# i=2
-# shap.force_plot(explainer.expected_value[0], shap_values[0][i], X_test.values[i], feature_names = X_test.columns)
-# 
-# 
-# fig.set_size_inches(18.5, 20.5)
-# fig.savefig("Z:\\R\\R projects\\LCA-GB\\models\\shap.png")
-# 
-# # -- first trial
-# shap.initjs()
-# shap_values = shap.TreeExplainer(model).shap_values(X_test)
-# shap.summary_plot(shap_values, X_test)
-# 
-# # -- second trial
-# scaler = StandardScaler()
-# scaled_data = scaler.fit_transform(X_test.iloc[:,:-1])
-# shap_values = explainer.shap_values(scaled_data.iloc[0,:-1].values.reshape(1,-1))
-# shap.force_plot(explainer.expected_value[0], shap_values[0], link='logit')  # repeat changing 0 for i in range(0, 5)
-# 
-# # -- third trial
-# shap.initjs()
-# class_names = ['Low', 'Moderate', ' High', 'Outstanding']
-# shap.summary_plot(shap_values, X_test, plot_type="bar", class_names= class_names, feature_names = ['Scenicness', 'Wildness'])
-# 
-# # -- plot the global importance of each feature
-# shap.plots.bar(explainer)
-# 
-# # visualize the first prediction's explanation
-# shap.force_plot(explainer.expected_value[0], shap_values[0])
-# 
-# # fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
-# # shap.summary_plot(shap_values)
-# # # or 
-# # shap.plots.beeswarm(shap_values)
-# shap.summary_plot(shap_values[2], X_test)
-# 
-# # fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
-# shap.summary_plot(shap_values, X_test, plot_type="bar")
-# 
-# shap_interaction_values = shap.TreeExplainer(model).shap_interaction_values(X_train)
-#     
-# X_names = ['Scenicness.median','x_coord','y_coord']
-# 
-# shap.summary_plot(shap_interaction_values, X_train, max_display=15, 
-#                   feature_names = X_names,
-#                   plot_type = 'dot')
-#                   # plot_type="compact_dot")
-#                   
-# import shap  # package used to calculate Shap values
-# 
-# # Create object that can calculate shap values
-# explainer = shap.TreeExplainer(my_model)
-# 
-# # calculate shap values. This is what we will plot.
-# # Calculate shap_values for all of val_X rather than a single row, to have more data for plot.
-# shap_values = explainer.shap_values(val_X)
-# 
-# # Make plot. Index of [1] is explained in text below.
-# shap.summary_plot(shap_values[1], val_X)                  
 ```
